### 2016-06-26 LinearRegression/Documentation/OneParam/OneParam.tex
adding first version of the documentation

\documentclass[12pt,journal]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{pgfplots}
\usepackage{caption}

\begin{document}

    \title{Lineal regression with one variable and one parameter for the cost
           function}
    \author{Alejandro Salgado G}
    \maketitle

    This problem consists in predicting differents ouputs based on some examples
    that the algorithm is going to receive. \\

    for example, we have the folowing points plotted in a graphic\\

    \begin{tikzpicture}
        \begin{axis}[ xmin=0, xmax=5,
                      ymin=0, ymax=10,
                      grid=both
                    ]
            \addplot[color=red, only marks]
                coordinates{
                    (1,2)
                    (1.5,1)
                    (1.5,3)
                    (2,2)
                    (2.5,4)
                    (3,5)
                    (3,4)
                    (3.5,4.5)
                    (3.5,5.5)
                    (4,5)
                    (4,6.5)
                    (4.5,8)
                };
        \end{axis}
    \end{tikzpicture}

    What the algorithm will do is to draw a line as closser as posible to all
    the points in order to get the best aproximation posible to the original
    answers\\

    \begin{tikzpicture}
        \begin{axis}[ xmin=0, xmax=5,
                      ymin=0, ymax=10,
                      grid=both
                    ]
            \addplot[color=blue]{1.5*x};
            \addplot[color=red, only marks]
                coordinates{
                    (1,2)
                    (1.5,1)
                    (1.5,3)
                    (2,2)
                    (2.5,4)
                    (3,5)
                    (3,4)
                    (3.5,4.5)
                    (3.5,5.5)
                    (4,5)
                    (4,6.5)
                    (4.5,8)
                };

        \end{axis}
    \end{tikzpicture}

    When the right line is founded we can get an aproximatly answers to
    inputs, that not necesarly where given in first place, allowing to
    get more info from data.\\

    \begin{tikzpicture}
        \begin{axis}[ xmin=0, xmax=5,
                      ymin=0, ymax=10,
                    ]
            \addplot[color=blue]{1.5*x};

            \draw[dashed] (axis cs: 0,1.125) -- node[above]{1.125} (axis cs: 0.75,1.125);
            \draw[dashed] (axis cs: 0.75,0) -- node[right]{0.75} (axis cs: 0.75,1.125);

            \draw[dashed] (axis cs: 0,2.25) -- node[above]{2.25} (axis cs: 1.5,2.25);
            \draw[dashed] (axis cs: 1.5,0) -- node[right]{1.5} (axis cs: 1.5,2.25);

            \draw[dashed] (axis cs: 0,3.375) -- node[above]{3.375} (axis cs: 2.25,3.375);
            \draw[dashed] (axis cs: 2.25,0) -- node[right]{2.25} (axis cs: 2.25,3.375);

            \draw[dashed] (axis cs: 0,4.6875) -- node[above]{4.6875} (axis cs: 3.125,4.6875);
            \draw[dashed] (axis cs: 3.125,0) -- node[right]{3.125} (axis cs: 3.125,4.6875);

            \draw[dashed] (axis cs: 0,6) -- node[above]{6} (axis cs: 4,6);
            \draw[dashed] (axis cs: 4,0) -- node[right]{4} (axis cs: 4,6);

            \addplot[color=blue, only marks]
                coordinates{
                    (0.75,1.125)
                    (1.5,2.25)
                    (2.25,3.375)
                    (3.125,4.6875)
                    (4,6)
                };

        \end{axis}
    \end{tikzpicture}

    In order to get this line we must define a function that give as the answers
    values, the functionwill be the equation of a line
    \begin{equation}
        h(x) = \theta_0 + \theta_1 x
    \end{equation}

    But taking into account that this is the first example that we are going to
    look at, we will eliminate the term that moves the graphic in the Y axis by
    make this variable equal to zero ($\theta_0 = 0$), this means  that the line
    that we will make will allways pass by the origin. In the next article this
    term will be used. In this article the function will be define as follow

    \begin{equation}
        h(x) = \theta x
    \end{equation}

    Now the problem will have two types of points: \\

    \begin{itemize}
        \item The result of the function $h(x)$
        \item The real result that we are going to call ($y$) \\
    \end{itemize}

    The next thing to do is compute how much the two sets of points differ, to
    do this we are going to use the cuadratic error of the two sets. Next we
    will give a practical example of how to get the formula. \\

    Lets say that in our problem we are going to have only three points to
    predict. Our function $h(x)$ gave the results $\{5,10,1\}$ and the correct
    results are $\{6,9,3\}$ and we want to cuantificate the error asocciated to
    this two answers.

    The first thing that must do is calculate the difference between this two
    sets:\\

    \begin{itemize}
        \item $5-6 = -1$
        \item $10-9 = 1$
        \item $1-3 = -2$ \\
    \end{itemize}

    The final result is another set $\{-1,1,-2\}$, this set contain positive and
    negative numbers, and if we decide to do the difference of the two sets in
    another order, for example results minus predictions, we will end up with a
    different set that can contain positive and negative numbers, as we only
    interested in calculate a number asociated to the error of our function
    $h(x)$ the negative numbers will not add any importat information, so we
    must eliminate this type of numbers in order to get only the information
    that we are interested on.

    To achive this gold we have two posible ways to do it, the first one is to
    add an absolute value to the calculation and the other is to squaring the
    results. The first option is not useful because later we will have to
    calculate a simple derivative to the resulting function and the absolute
    value is not derivable, this let us with the suquare option\\

    \begin{itemize}
        \item $(5-6)^2 = 1$
        \item $(10-9)^2 = 1$
        \item $(1-3)^2 = 4$\\
    \end{itemize}

    The result of this operation is the set $\{1,1,4\}$. Exelent, now we will
    never have a negative number in out resulting set, notice that our
    answers will be much bigger because we are using square of two as our
    method to eliminate the negative numbers, actually this doesn't metter
    because we are lookin for a number that describes the error associated to
    the prediction that we are computing so the range of the number is not too
    important.\\

    Now that we have our resulting set we want to get the mean of the errors,
    and this will be our error associated to the prediction that we have done\\

    \begin{itemize}
        \item $\frac{1+1+4}{3} = \frac{6}{3} = 2$ \\
    \end{itemize}

    Now that we have seen the process, we are gonig to formalize a little bit
    the process. The first thing that we are going to do is define a varible as
    the number of elements in each set

    \begin{itemize}
        \item $m = number\_of\_elements$
    \end{itemize}

    Second we are going to define the resulting set as $\{x_1,x_2, ... , x_m\}$
    and the real resulting set as $\{y_1,y_2, ... , y_m\}$. This definition can
    be more concise because the first set is the result of the function $h(x)$,
    so we are going to define it as $h(x_i)$ and the second can be decribed as
    $y_i$.

    Now we will define the substration of each set as follows

    \begin{equation}
        h(x_i) - y_i \hspace{0.5cm} i=1, 2, ..., m
    \end{equation}

    Then adding the power of two part we will get the next expession

    \begin{equation}
        (h(x_i) - y_i)^2 \hspace{0.5cm} i=1, 2, ..., m
    \end{equation}

    Now adding the mean calculation we get

    \begin{equation}
        \frac{\sum_{i=1}^{m} (h(x_i) - y_i)^2}{m}
    \end{equation}

    Applying properties of sum

    \begin{equation}
       \frac{1}{m} \sum_{i=1}^{m} ( h(x_i) - y_i )^2
    \end{equation}

    Finally to make easier the future derivation we will add another term, a
    division by to, this step will be explained later.

    \begin{equation}
       \frac{1}{2m} \sum_{i=1}^{m} ( h(x_i) - y_i )^2
    \end{equation}

    This is the cuadratic error and will be used as our cost function in terms
    of the parameter $\theta$ to expess the error associated to the prediction
    of the algorithm with respect to the correct output that is given at the
    begin of the problem.

    \begin{equation}
       J(\theta) = \frac{1}{2m} \sum_{i=1}^{m} ( h(x_i) - y_i )^2
    \end{equation}

    This equation has the following graphic

    \begin{figure}[h!]

        \centering
        \captionsetup{justification=centering}

        \begin{tikzpicture}
            \begin{axis}[
                            xlabel=$\theta$,
                            ylabel=$J(\theta)$,
                            xmin=-5, xmax=5,
                            ymin=0, ymax=10,
                        ]
                \addplot[mark=none, color=blue]{x^2};
            \end{axis}
        \end{tikzpicture}

        \caption{Cost function}
    \end{figure}

    The next step is define a way to make the error as small as posible,
    to do this we will use a famus algorithm called gradient decend.
    the formula of the algorithm is show bellow

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{d}{d \theta} J(\theta)
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{d}{d \theta}
                    \Bigg[
                        \frac{1}{2m} \sum_{i=1}^{m} ( h(x_i) - y_i )^2
                    \Bigg]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha \frac{1}{2m}
            \Bigg(
                \frac{d}{d \theta}
                    \Bigg[
                        \sum_{i=1}^{m} (\theta x_i - y_i)^2
                    \Bigg]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{1}{2m} \sum_{i=1}^{m}
                    \Bigg[
                        \frac{d}{d \theta} (\theta x_i - y_i)^2
                    \Bigg]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{1}{2m} \sum_{i=1}^{m}
                    \Bigg[
                        2(\theta x_i - y_i) \frac{d}{d \theta}(\theta x_i - y_i)
                    \Bigg]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{1}{2m} \sum_{i=1}^{m}
                    \Big[
                        2 (\theta x_i - y_i) (x_i - 0)
                    \Big]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{1}{2m} \sum_{i=1}^{m}
                    \Big[
                        2 (\theta x_i - y_i) x_i
                    \Big]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{2}{2m} \sum_{i=1}^{m}
                    \Big[
                        (\theta x_i - y_i) x_i
                    \Big]
            \Bigg)
    \end{equation}

    \begin{equation}
        \theta := \theta - \alpha
            \Bigg(
                \frac{1}{m} \sum_{i=1}^{m}
                    \Big[
                        (\theta x_i - y_i) x_i
                    \Big]
            \Bigg)
    \end{equation}

\end{document}

### 2016-06-26 LinearRegression/Documentation/OneParam/OneParam.tex
adding gradient decend explanation for lineal regression with one varible

@@
 
+    The objective of this function is to actualize the parameter $\theta$
+    of the cost function J minimizing the error in each iteration, this
+    is done by searching the path to the global minimum. The function
+    that we are going to use is described bellow.\\
+
+    The first thing to do is replace the function $J(\theta)$
+
     \begin{equation}
@@
     \end{equation}
+
+    Now the factor of the denominator has no influence on the function so it
+    can be tranlated outside of the derivative
 
@@
 
+    The derivative of the summatory is equal to the derivative of each term and
+    then add each result, so we can insert the dirivative into the summary
+
     \begin{equation}
@@
     \end{equation}
+
+    Then we calculate the derivative, notice that because of derivative
+    properties we must calculate the outside derivative multiplied by the inside
+    derivative
 
@@
 
+    Now we calculate the inside derivative
+
     \begin{equation}
@@
 
-    \begin{equation}
-        \theta := \theta - \alpha
-            \Bigg(
-                \frac{1}{2m} \sum_{i=1}^{m}
-                    \Big[
-                        2 (\theta x_i - y_i) x_i
-                    \Big]
-            \Bigg)
-    \end{equation}
+    Now we can take the factor 2 outside the summatory
 
@@
     \end{equation}
+
+    Finally we can see the reason why the denominator has a extra factor
+    (the 2 that we added at the begining), this is with the objective
+    to cancel the 2 in the numerator that result of the derivative. Finaly
+    this is the resulting function that we are going to use to minimize the
+    error of the algorithm
 

### 2016-06-26 LinearRegression/Documentation/OneParam/OneParam.tex
adding gradient decend plots

@@
     \begin{tikzpicture}
-        \begin{axis}[ xmin=0, xmax=5,
+        \begin{axis}[
+                      title={Input points},
+                      xlabel=$x$,
+                      ylabel=$y$,
+                      xmin=0, xmax=5,
                       ymin=0, ymax=10,
@@
     \begin{tikzpicture}
-        \begin{axis}[ xmin=0, xmax=5,
+        \begin{axis}[
+                      xmin=0, xmax=5,
                       ymin=0, ymax=10,
@@
     \begin{tikzpicture}
-        \begin{axis}[ xmin=0, xmax=5,
+        \begin{axis}[
+                      title={Predictions},
+                      xlabel=$h(x)$,
+                      ylabel=$y$,
+                      xmin=0, xmax=5,
                       ymin=0, ymax=10,
@@
     the process. The first thing that we are going to do is define a varible as
-    the number of elements in each set
+    the number of elements in each set \\
 
     \begin{itemize}
-        \item $m = number\_of\_elements$
+        \item $m = number\_of\_elements$ \\
     \end{itemize}
@@
 
-    This equation has the following graphic
+    This equation will have the following graphic
 
     \begin{figure}[h!]
-
-        \centering
-        \captionsetup{justification=centering}
-
         \begin{tikzpicture}
             \begin{axis}[
+                         title={Cost function},
                          xlabel=$\theta$,
                          ylabel=$J(\theta)$,
-                            xmin=-5, xmax=5,
-                            ymin=0, ymax=10,
+                         ymin=0, ymax=25,
+                         xtick=\empty,
+                         ytick=\empty
                         ]
@@
         \end{tikzpicture}
-
-        \caption{Cost function}
     \end{figure}
@@
     of the cost function J minimizing the error in each iteration, this
-    is done by searching the path to the global minimum. The function
-    that we are going to use is described bellow.\\
+    is done by searching the path to the global minimum. \\
 
-    The first thing to do is replace the function $J(\theta)$
+    \begin{figure}[ht!]
+        \begin{tikzpicture}
+            \begin{axis}[
+                         xlabel=$\theta$,
+                         ylabel=$J(\theta)$,
+                         ymin=0, ymax=25,
+                         xtick=\empty,
+                         ytick=\empty
+                        ]
+                \addplot[mark=none, color=red]{x^2};
+
+                \draw[thick,->] (axis cs: 3.5,16) -- node[left]{$\alpha \Bigg(\frac{d}{d \theta} J(\theta)\Bigg)$} (axis cs: 2.65,9.75);
+                \draw[thick,->] (axis cs: 2.5,9) -- (axis cs: 1.65,4.75);
+                \draw[thick,->] (axis cs: 1.5,4) -- (axis cs: 0.75,2);
+                \draw[thick,->] (axis cs: 0.75,1.75) -- (axis cs: 0,0.75);
+
+                \addplot[color=blue, only marks]
+                coordinates{
+                    (4,16)
+                    (3,9)
+                    (2,4)
+                    (1,1)
+                    (0,0)
+                };
+            \end{axis}
+        \end{tikzpicture}
+    \end{figure}
+
+    \begin{figure}[ht!]
+        \begin{tikzpicture}
+            \begin{axis}[
+                         title={Gradient decend with negative derivative},
+                         xlabel=$\theta$,
+                         ylabel=$J(\theta)$,
+                         ymin=0, ymax=25,
+                         xtick=\empty,
+                         ytick=\empty
+                        ]
+                \addplot[mark=none, color=red]{x^2};
+
+                \draw[thick,->] (axis cs: -3.5,16) -- node[right]{$\alpha\Bigg(\frac{d}{d\theta}J(\theta)\Bigg)$} (axis cs: -2.65,9.75);
+                \draw[thick,->] (axis cs: -2.5,9) -- (axis cs: -1.65,4.75);
+                \draw[thick,->] (axis cs: -1.5,4) -- (axis cs: -0.75,2);
+                \draw[thick,->] (axis cs: -0.75,1.75) -- (axis cs: -0,0.75);
+
+                \addplot[color=blue, only marks]
+                coordinates{
+                    (-4,16)
+                    (-3,9)
+                    (-2,4)
+                    (-1,1)
+                    (0,0)
+                };
+            \end{axis}
+        \end{tikzpicture}
+    \end{figure}
+
+    The function that we are going to use is described bellow. The first thing
+    to do is replace the function $J(\theta)$
 

### 2016-06-27 LinearRegression/Documentation/Derivative/Derivative.tex
starting the derivative explanation document

\documentclass[12pt,journal]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{pgfplots}
\usepackage{caption}

\begin{document}

    \title{Derivative fundamentals}
    \author{Alejandro Salgado Gomez}

    \maketitle

    This article will try to explain as simple as posible the concept of a
    derivative. A derivatavie is the solution to one of the principal
    problems in calculus, which is how to calculate the tangent rect throught
    a point, in other words, how to calculate the rate of change of a function.

    This has a lot of practical applications in mathematics and the real
    world\\

    Lets start from the beginning, lets say we have two points that represent
    the state of something that is moving, the first one is the starting
    position and the second one is the final position. And we want to calculate
    how much is the change of the possition of this thing from the starting
    point to the final position. Here a ilustration

    \begin{tikzpicture}
        \begin{axis}[
                      xlabel=$x$,
                      ylabel=$y$,
                      xmin=0, xmax=10,
                      ymin=0, ymax=10,
                      grid=both
                    ]
            \addplot[color=red, only marks]
                coordinates{
                    (2,2)
                };

            \addplot[color=blue, only marks]
                coordinates{
                    (4,6)
                };

            \draw[thick,->] (axis cs: 2.1,2.2) -- node[above]{?} (axis cs: 3.9,5.8);
        \end{axis}
    \end{tikzpicture}

    Notice that we are not interested in calculate the distance between this
    points, what we want is find a number that decribes how much the thing's
    position has change when he moves from the beginning to the end.

    This can be done by calculating for each unit of distance that the object
    has move, how much the object assend. This calculus can be  described as
    follows

    \begin{equation}
        change = \frac{ascend}{advance}
    \end{equation}

\end{document}

### 2016-06-28 LinearRegression/Documentation/Derivative/Derivative.tex
defining the derivative

@@
     the state of something that is moving, the first one is the starting
-    position and the second one is the final position. And we want to calculate
-    how much is the change of the possition of this thing from the starting
-    point to the final position. Here a ilustration
+    position and the second one is the position after a slide of time. And we
+    want to calculate how much is the change of the possition of this thing
+    from the starting point to the final position. Here a ilustration\\
 
@@
                       ylabel=$y$,
-                      xmin=0, xmax=10,
-                      ymin=0, ymax=10,
+                      xmin=0, xmax=6,
+                      ymin=0, ymax=8,
                       grid=both
@@
 
+            \node[below] at (axis cs: 2,2) {(2,2)};
+
             \addplot[color=blue, only marks]
@@
                 };
+
+            \node[above] at (axis cs: 4,6) {(4,6)};
 
@@
     Notice that we are not interested in calculate the distance between this
-    points, what we want is find a number that decribes how much the thing's
-    position has change when he moves from the beginning to the end.
+    points, what we want to find is a number that decribes how much the thing's
+    position has change when he moves from the beginning to the second possition.
 
@@
 
+    \vspace{1cm}
+
+    For the example the formula would be $change = \frac{2}{1} = 2$, this
+    means that every time that the thing moves 1 unit, also assend two units.
+    Now the illustration \\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x$,
+                      ylabel=$y$,
+                      xmin=0, xmax=6,
+                      ymin=0, ymax=8,
+                      grid=both
+                    ]
+            \addplot[color=red, only marks]
+                coordinates{
+                    (2,2)
+                };
+
+            \addplot[color=black, only marks]
+                coordinates{
+                    (3,4)
+                };
+
+            \addplot[color=blue, only marks]
+                coordinates{
+                    (4,6)
+                };
+
+            \draw[dashed,->] (axis cs: 2.1,2.2) -- (axis cs: 3.9,5.8);
+
+            \draw[thick,->] (axis cs: 2,2) -- node[below]{1} (axis cs: 3,2);
+            \draw[thick,->] (axis cs: 3,2) -- node[left]{2} (axis cs: 3,4);
+
+            \draw[thick,->] (axis cs: 3,4) -- node[below]{1} (axis cs: 4,4);
+            \draw[thick,->] (axis cs: 4,4) -- node[left]{2} (axis cs: 4,6);
+        \end{axis}
+    \end{tikzpicture}
+
+    Now that we know the "rate of change" of the movement we can calculate which
+    would be the position of the object at any moment of the movement, this means
+    that we can know the position when the object has move 1 unit, 2 units, or
+    any amount of units that we want.
+
+    \begin{figure}[h]
+        \begin{tikzpicture}
+            \begin{axis}[
+                          xlabel=$x$,
+                          ylabel=$y$,
+                          xmin=0, xmax=10,
+                          ymin=0, ymax=20,
+                          grid=both
+                        ]
+                \addplot[color=red, only marks]
+                    coordinates{
+                        (2,2)
+                        (4,6)
+                    };
+
+                \addplot[color=blue, only marks]
+                    coordinates{
+                        (6,10)
+                        (8,14)
+                    };
+
+                \addplot[color=black, only marks]
+                    coordinates{
+                        (6,2)
+                        (8,2)
+                    };
+
+                \draw[thick,->] (axis cs: 2,2) -- node[below]{2} (axis cs: 4,2);
+                \draw[thick,->] (axis cs: 4,2) -- node[right]{4} (axis cs: 4,5.6);
+
+                \draw[dashed,->] (axis cs: 4,2) -- node[below]{4} (axis cs: 5.8,2);
+                \draw[dashed,->] (axis cs: 6,2) -- node[right]{8} (axis cs: 6,9.6);
+
+                \draw[dashed,->] (axis cs: 6,2) -- node[below]{6} (axis cs: 7.8,2);
+                \draw[dashed,->] (axis cs: 8,2) -- node[right]{12} (axis cs: 8,13.6);
+
+                \draw[dashed,->] (axis cs: 2,2) -- (axis cs: 7.8,13.6);
+            \end{axis}
+        \end{tikzpicture}
+    \end{figure}
+
+    This means that now we can predict how is going to be the movement of the
+    object. All we have to do is take the units of movement that we want in the
+    horizontal axis and then multiply that number for the change that we found
+    before and thats all.\\
+
+    Now we are going to formalize this concept a little bit. First we are going
+    to change the name of the variable "change" to just a letter for simplicity,
+    lets say "m". \\
+
+    $change = m$ \\
+
+    Now we are going to take the define the horizontal distance between the points
+    as the difference between the two $x$ components and the vertical distance as
+    the difference between the two $y$ components. In mathematics the difference
+    between a same variable is denoted as $\Delta$variable, this means that the
+    difference between the $x$ component will be $\Delta x$ and the difference
+    between the $y$ components will be $\Delta y$\\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x$,
+                      ylabel=$y$,
+                      xmin=0, xmax=7,
+                      ymin=0, ymax=8,
+                      xtick=\empty,
+                      ytick=\empty,
+                    ]
+            \addplot[color=red, only marks]
+                coordinates{
+                    (2,2)
+                };
+            \node[above] at (axis cs: 2,2) {($x_1$,$y_1$)};
+
+            \addplot[color=blue, only marks]
+                coordinates{
+                    (4,6)
+                };
+            \node[left] at (axis cs: 4,6) {($x_2$,$y_2$)};
+
+            \draw[thick] (axis cs: 2,2) -- node[below]{$x_1 - x_2 = \Delta x$}(axis cs: 4,2);
+            \draw[thick] (axis cs: 4,2) -- node[right]{$y_1 - y_2 = \Delta y$} (axis cs: 4,6);
+        \end{axis}
+    \end{tikzpicture}
+
+    Then we can define the change as follows
+
+    \begin{equation}
+        m = \frac{\Delta y}{\Delta x}
+    \end{equation}
+
+    The last formula is the formal definition of the slope in mathematics, this
+    is the change of state in a point of a linear function, just like we have
+    explained earlier.
+
+    This works very good for any linear function, but what happens when we
+    whant to analise a non linear function like a curve?. Lets see an example
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x$,
+                      ylabel=$y$,
+                      xmin=0, xmax=6,
+                      ytick=\empty
+                    ]
+
+            \addplot[color=blue]{x^4};
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (1,1)
+                };
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (4.8,530)
+                };
+
+            \addplot[color=black]{139.2*x -138.2};
+        \end{axis}
+    \end{tikzpicture}
+
 \end{document}

### 2016-06-29 LinearRegression/Documentation/Derivative/Derivative.tex
all derivative information ready

@@
     This article will try to explain as simple as posible the concept of a
-    derivative. A derivatavie is the solution to one of the principal
-    problems in calculus, which is how to calculate the tangent rect throught
-    a point, in other words, how to calculate the rate of change of a function.
-
-    This has a lot of practical applications in mathematics and the real
-    world\\
+    derivative. A derivativie is the solution to one of the principal
+    problems in calculus, which is how to calculate the slope of the tangent
+    rect throught a point, in other words, how to calculate the rate of change
+    of a function, don't worry if this concept is confusing, we are going to
+    give varouis examples to ilustrate this ideas.
 
@@
         \begin{axis}[
-                      xlabel=$x$,
-                      ylabel=$y$,
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
                       xmin=0, xmax=6,
@@
     Now that we know the "rate of change" of the movement we can calculate which
-    would be the position of the object at any moment of the movement, this means
-    that we can know the position when the object has move 1 unit, 2 units, or
-    any amount of units that we want.
+    would be the position of the object at any moment of the movement, this
+    means that we can know the position when the object has move 1 unit, 2
+    units, or any amount of units that we want.
 
-    \begin{figure}[h]
+    \begin{figure}[h!]
         \begin{tikzpicture}
             \begin{axis}[
-                          xlabel=$x$,
-                          ylabel=$y$,
+                          xlabel=$x-axis$,
+                          ylabel=$y-axis$,
                           xmin=0, xmax=10,
@@
 
-    Now we are going to take the define the horizontal distance between the points
+    Now we are going to define the horizontal distance between the points
     as the difference between the two $x$ components and the vertical distance as
@@
         \begin{axis}[
-                      xlabel=$x$,
-                      ylabel=$y$,
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
                       xmin=0, xmax=7,
@@
 
-            \draw[thick] (axis cs: 2,2) -- node[below]{$x_1 - x_2 = \Delta x$}(axis cs: 4,2);
-            \draw[thick] (axis cs: 4,2) -- node[right]{$y_1 - y_2 = \Delta y$} (axis cs: 4,6);
+            \draw[thick] (axis cs: 2,2) -- node[below]{$x_2 - x_1 = \Delta x$}(axis cs: 4,2);
+            \draw[thick] (axis cs: 4,2) -- node[right]{$y_2 - y_1 = \Delta y$} (axis cs: 4,6);
         \end{axis}
@@
     This works very good for any linear function, but what happens when we
-    whant to analise a non linear function like a curve?. Lets see an example
+    want to analise a non linear function like a curve?. Lets analize an
+    example, we are goint to take a fuction with a curve graphic, we are going to
+    take a point and we want to calculate the change in that specific point,
+    this can be achived by finding the value of the slope in that point, in other
+    words finding the rate of change in that specific point.
+
+    The problem is that we only know how to calculate the slope with two points,
+    so we are goint to take a second point with reference to the original point,
+    this difference is goint to be a number called "$h$" \\
 
@@
         \begin{axis}[
-                      xlabel=$x$,
-                      ylabel=$y$,
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
                       xmin=0, xmax=6,
+                      ymin=-200,
+                      xtick=\empty,
                       ytick=\empty
@@
             \addplot[color=blue]{x^4};
+            \addlegendentry{$f(x)$}
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (1,1)
+                };
+            \node[above] at (axis cs: 1,1) {($x$,$y_1$)};
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (4.8,530)
+                };
+            \node[left] at (axis cs: 4.8,530) {($x+h$,$y_2$)};
+
+            \draw (axis cs: 1,-200) -- (axis cs: 1,1);
+            \draw (axis cs: 4.8,-200) -- (axis cs: 4.8,530);
+            \draw (axis cs: 1,-100) -- node[above]{$h$} (axis cs: 4.8,-100);
+        \end{axis}
+    \end{tikzpicture}
+
+    Notice that those points are part of a function, this means that the high
+    of any point in the graphic can be calculated as the function evaluated in
+    the point of the $x$ axis that we want to evaluate. Now an example \\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
+                      xmin=0, xmax=6,
+                      ymin=-200,
+                      xtick=\empty,
+                      ytick=\empty
+                    ]
+
+            \addplot[color=blue]{x^4};
+            \addlegendentry{$f(x)$}
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (1,1)
+                };
+            \node[above] at (axis cs: 1,1) {($x$,$f(x)$)};
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (4.8,530)
+                };
+            \node[left] at (axis cs: 4.8,530) {($x+h$,$f(x +h)$)};
+
+            \draw (axis cs: 1,-200) -- node[right]{$f(x)$} (axis cs: 1,1);
+            \draw (axis cs: 4.8,-200) -- (axis cs: 4.8,530);
+
+            \node[left] at (axis cs: 4.8,-100) {$f(x+h)$};
+        \end{axis}
+    \end{tikzpicture}
+
+    Now with this new concepts we can define the slope of this two points as
+
+    \begin{equation}
+        m = \frac{f(x+h) - f(x)}{(x+h) - x}
+    \end{equation}
+
+    We can simplify the last equation
+
+    \begin{equation}
+        m = \frac{f(x+h) - f(x)}{h}
+    \end{equation}
+
+    Now with this new formula we can calculate the slope of the two points and
+    find the line that cross the points.
+
+    Regrettably this slope does not describe the change of the point very well
+    because the line does not touch the function in just one point (this is
+    what is a tangent line), insted touch the function two times (this is call a 
+    secant line), so this is not what we want. Now the ilustration \\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
+                      xmin=0, xmax=6,
+                      ymin=-100, ymax=800,
+                      xtick=\empty,
+                    ]
+
+            \addplot[color=blue]{x^4};
+            \addlegendentry{$f(x)$}
+
+            \addplot[color=red]{171.5*x-450};
+            \addlegendentry{$tangent$}
+
+            \addplot[color=black]{157.6*x-299};
+            \addlegendentry{$secant$}
+
+            \addplot[color=black, only marks]
+                coordinates{
+                    (3.5,150)
+                    (2,16)
+                    (4.5,410)
+                };
+
+        \end{axis}
+    \end{tikzpicture}
+
+    Until this moment we have the following graphic \\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
+                      xmin=0, xmax=6,
+                      ymin=-200, ymax=850,
+                      xtick=\empty,
+                      ytick=\empty
+                    ]
+
+            \addplot[color=blue]{x^4};
+            \addlegendentry{$f(x)$}
+
+            \addplot[color=red]{139.2*x-138};
+            \addlegendentry{rate of change}
 
@@
 
-            \addplot[color=red, only marks]
+            \addplot[color=blue, only marks]
                 coordinates{
@@
 
-            \addplot[color=black]{139.2*x -138.2};
+            \draw (axis cs: 1,-200) -- (axis cs: 1,1);
+            \draw (axis cs: 4.8,-200) -- (axis cs: 4.8,530);
+            \draw (axis cs: 1,-100) -- node[above]{$h$} (axis cs: 4.8,-100);
         \end{axis}
@@
 
+    What happens if we start to make $h$ smaller?, the points will be closser
+    every time, so the linear function is going to start be much more closer
+    to predict the correct rate of change of the point that we are analyzing,
+    in other words the secant line that we have will tend to be a tangent, just
+    as we want. Here the ilustration
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
+                      xmin=0, xmax=6,
+                      ymin=-200, ymax=850,
+                      xtick=\empty,
+                      ytick=\empty
+                    ]
+
+            \addplot[color=blue]{x^4};
+            \addlegendentry{$f(x)$}
+
+            \addplot[color=black]{139.2*x-138};
+            \addlegendentry{rates of change}
+
+            \addplot[color=black]{85*x-84};
+            \addplot[color=black]{40*x-39};
+            \addplot[color=black]{15*x-14};
+
+            \addplot[color=red, only marks]
+                coordinates{
+                    (1,1)
+                };
+
+            \addplot[color=blue, only marks]
+                coordinates{
+                    (4.8,530)
+                    (4,256)
+                    (3,81)
+                    (2,16)
+                };
+
+            \draw (axis cs: 1,-200) -- (axis cs: 1,1);
+            \draw (axis cs: 4.8,-200) -- (axis cs: 4.8,530);
+            \draw (axis cs: 4,-200) -- (axis cs: 4,256);
+            \draw (axis cs: 3,-200) -- (axis cs: 3,81);
+            \draw (axis cs: 2,-200) -- (axis cs: 2,16);
+
+            \draw[thick, color=red] (axis cs: 1,-100) -- node[below]{$h$} (axis cs: 4.8,-100);
+        \end{axis}
+    \end{tikzpicture}
+
+    Is for this reason that the derivative is define as follows
+
+    \begin{equation}
+        \lim_{h \to 0} \frac{f(x+h) - f(x)}{h}
+    \end{equation}
+
+    Finally this is the formal definition of a derivative, if we apply this
+    concept to the previus graphics we can get directly the tanget line that
+    pass through a specific point of a function and the slope or rate of change
+    of that point. \\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                      xlabel=$x-axis$,
+                      ylabel=$y-axis$,
+                      xmin=0, xmax=6,
+                      ymin=-200, ymax=850,
+                      xtick=\empty,
+                      ytick=\empty
+                    ]
+
+            \addplot[color=blue]{x^4};
+            \addlegendentry{$f(x)$}
+
+            \addplot[color=red]{4*x-3};
+            \addlegendentry{$Derivative$}
+
+            \addplot[color=red]{171.5*x-450};
+            \addplot[color=red]{415.3*x-1468};
+
+            \addplot[color=black, only marks]
+                coordinates{
+                    (1,1)
+                    (3.5,150)
+                    (4.7,484)
+                };
+        \end{axis}
+    \end{tikzpicture}
+
+    As we can see in the last graphic the slope of the first point is almost 
+    horizontal, this means that the rate of change is samall, in the second
+    is bigger and in the third point is the most vertical, so is the point with
+    the highest change in the graphic, just like it would be, so we can
+    conclude that with the derivative we can know the rate of change the
+    functions at any point (almost, see defferentiable functions for more
+    information).
+
 \end{document}

### 2016-06-29 LinearRegression/Documentation/OneParam/OneParam.tex
starting gradient decendent explanation

@@
     of the cost function J minimizing the error in each iteration, this
-    is done by searching the path to the global minimum. \\
+    is done by searching the path to the global minimum.
+
+    for those of you who are not familiar with the derivarives, is highly
+    recomendable to look at the Derivative fundamentals article that is in
+    located in the documentation directory.\\
+
+    Now lets analyze the formula, the fist simbol is the $\theta$ that is the
+    parameter of the cost function, the second is the $:=$ which means assign,
+    this make sence because we are trying to update the parameter of the
+    function trying to make it as closer to the minimum as possible.
 
@@
             \begin{axis}[
+                         title={Gradient decend with positive derivative},
                          xlabel=$\theta$,
@@
 
-    The function that we are going to use is described bellow. The first thing
-    to do is replace the function $J(\theta)$
+    \vspace{8cm}
+
+    Now the first thing to do in order to use the gradient desent in our cost
+    function is replace the function $J(\theta)$ in the formula
 

### 2016-06-29 LinearRegression/Documentation/Derivative/Derivative.tex
adding more detail in the tangent vs secant part

@@
     because the line does not touch the function in just one point (this is
-    what is a tangent line), insted touch the function two times (this is call a 
-    secant line), so this is not what we want. Now the ilustration \\
+    what a tangent line is), insted touch the function two times (this is call a
+    secant line), this means that the change of the point that we are
+    calculating is taking into account two points of the function, not only the
+    original point, so this is not what we want. Now the ilustration \\
 
@@
                       xtick=\empty,
+                      ytick=\empty,
                     ]

### 2016-06-29 LinearRegression/Documentation/OneParam/OneParam.tex
adding joke to the oneParam document

@@
 \documentclass[12pt,journal]{IEEEtran}
+
+%
 
@@
 \usepackage{caption}
+\usepackage{graphicx}
+\graphicspath{ {Images/} }
 
@@
 
-    for those of you who are not familiar with the derivarives, is highly
+    for those of you who are not familiar with the derivatives, is highly
     recomendable to look at the Derivative fundamentals article that is in
     located in the documentation directory.\\
+
+    Before explaining the deeper meaning of this derivative we are going to
+    explain in other words what the gradient decend does, imagin that you are
+    lost in a jungle that is kind of a valley, and you are trying to find your
+    frinds that are in the lowest part of the valley, but you dont know exactly
+    in what point of the valley are you, what do you do?\\
+
+    %taken from http://bluepoppyholidays.com/wp-content/uploads/2014/10/Valley-of-flowers-from-top.jpg
+
+    \vspace{2cm}
+
+    \begin{figure}[h]
+        \centering
+        \includegraphics[scale=0.2]{Valley}
+        \caption{What you are specting as a valley}
+    \end{figure}
+
+    \begin{figure}[h]
+        \centering
+        \begin{tikzpicture}
+            \begin{axis}[
+                         title=What you have to imagin as a valley,
+                         ymin=0, ymax=25,
+                         xtick=\empty,
+                         ytick=\empty
+                        ]
+                \addplot[mark=none, color=black]{x^2};
+
+                \node[above] at (axis cs: 0,2) {Your};
+                \node[above] at (axis cs: 0,0.5) {friends};
+
+                \addplot[color=blue, only marks]
+                    coordinates{
+                        (0,0)
+                    };
+
+                \addplot[color=red, only marks]
+                    coordinates{
+                        (4,16)
+                        (2,4)
+                        (-3,9)
+                    };
+
+                \addplot[color=black, only marks]
+                    coordinates{
+                        (0,18)
+                    };
+
+                \node[above] at (axis cs: 0,18) {Your possible possition};
+
+                \draw[thick, ->] (axis cs: 0,18) -- (axis cs: 3.7,16);
+                \draw[thick, ->] (axis cs: 0,18) -- (axis cs: 1.8,4.4);
+                \draw[thick, ->] (axis cs: 0,18) -- (axis cs: -2.9,9.3);
+            \end{axis}
+        \end{tikzpicture}
+    \end{figure}
+
+    What you shoul do
 
@@
     this make sence because we are trying to update the parameter of the
-    function trying to make it as closer to the minimum as possible.
+    function trying to make it as closer to the minimum as possible. Then we
+    have a the actual value of $\theta$ followed by a substraction and a
+    $\alpha$ simbol, we are goint to talk about this later, now we get to de
+    most important part of the formula, the derivative ($\frac{d}{d \theta}$).\\
 
-    \begin{figure}[ht!]
+    the idea is that we are going to start in a point that is located at any
+    part of the graphic, it could be far or close to the minimum, remember that
+    this point is the result of the cost function for a line, and what we want
+    to do is to start updating the value of $\theta$ because is the value that
+    makes that the error will be big or small ($J(\theta)$).\\
+
+    \begin{tikzpicture}
+        \begin{axis}[
+                     xlabel=$\theta$,
+                     ylabel=$J(\theta)$,
+                     ymin=0, ymax=25,
+                     xtick=\empty,
+                     ytick=\empty
+                    ]
+            \addplot[mark=none, color=red]{x^2};
+
+            \draw[thick,->] (axis cs: 4,0) -- node[left]{$\theta_1$} (axis cs: 4,16);
+
+            \addplot[color=blue, only marks]
+            coordinates{
+                (4,16)
+                (2,4)
+                (1,1)
+                (0,0)
+            };
+        \end{axis}
+    \end{tikzpicture}
+
+    This can be done by calculating the direcction and that we want to follow
+
     \begin{tikzpicture}
@@
     \end{tikzpicture}
-    \end{figure}
 
-    \begin{figure}[ht!]
     \begin{tikzpicture}
@@
     \end{tikzpicture}
-    \end{figure}
-
-    \vspace{8cm}
 

### 2016-07-08 OctaveExamples/Global/Global.m
adding global variable example in octave

@@ -1,0 +1,8 @@
+global ar = 1;
+
+function ret = test()
+    global ar;
+    ret = ar + 1;
+endfunction
+
+test()

### 2016-07-08 LinearRegression/Documentation/OneParam/OneParam.tex
adding octave implementation all, info with one paramter redy

@@
 \documentclass[12pt,journal]{IEEEtran}
-
-%
 
@@
 
-    What you shoul do
+    Lets see how can we make our way to the bottom of the valley, the first
+    thing to do is find a direction, in order to do this we must know the
+    inclination of the side where you are, then start going to the lower place.
+    Lets assume that you start in the top left point, when you start going down
+    the hill, with a constant step size (this is very important for the
+    example), you should notice that you have a big advance with each step
+    becouse of the high inclination, but when you are aproaching your friends
+    the inclunation is lower and you start to get smaller steps until you get
+    your friends and stop walking.
 
-    Now lets analyze the formula, the fist simbol is the $\theta$ that is the
-    parameter of the cost function, the second is the $:=$ which means assign,
-    this make sence because we are trying to update the parameter of the
-    function trying to make it as closer to the minimum as possible. Then we
-    have a the actual value of $\theta$ followed by a substraction and a
-    $\alpha$ simbol, we are goint to talk about this later, now we get to de
-    most important part of the formula, the derivative ($\frac{d}{d \theta}$).\\
+    \newpage
 
-    the idea is that we are going to start in a point that is located at any
-    part of the graphic, it could be far or close to the minimum, remember that
-    this point is the result of the cost function for a line, and what we want
-    to do is to start updating the value of $\theta$ because is the value that
-    makes that the error will be big or small ($J(\theta)$).\\
+    Now that we have seen the example lets analyze the formula, the fist symbol
+    is the $\theta$ that in the example is your position but in the
+    formula is the parameter of the cost function, the second symbol is $:=$
+    which means that we are updating the value or assigning a new one,
+    this make sence because we are trying to get closer to our friends which
+    means update the cost function ($J(\theta)$) to the minimum. Then we
+    have a second $\theta$ that is your actual possition in the example and
+    the actual value of $\theta$ in the formula, followed by a substraction
+    that means that we are going to decrement the actual value of $\theta$
+    and then assign the new value to $\theta$ trying to get the minimum,
+    that means update the value, in the example means walk, then we get an
+    $\alpha$ symbol which is a positive number, this represent the step size,
+    and is a constant, now we get to the most important part of the formula, the
+    derivative ($\frac{d}{d \theta}$).
 
+    Remember that the first thing we do in the example was looked at the
+    inclunation of the hill where we are, this in order to know if we should
+    move right or left. Now, as you may know a derivative is basicaly the
+    slope of a function in a point, this is perfect for us because fit our
+    request perfectly. Lets imaging that we start at the right side of the
+    valley, so we what to go left, now mathematically the derivative of the
+    point situated in the right side of the function is positive because the
+    slope is pointing the right, so the term $\frac{d}{d \theta}$ will be a
+    positive number, that will be multiplied by a factor ($\alpha$) and then
+    substracted to the actual value of $\theta$, just as we want.
+    Now lets imaging that we end up in the left side of the valley, in that case
+    we want to go right, now mathematically the derivative of a point that is
+    situated in the left side of the function will be negative because the
+    slope is pointing to the left, this means that the term $\frac{d}{d \theta}$
+    will be a negative number, that will be multiplied by the factor ($\alpha$)
+    and then added to the actual value of $\theta$ (remember that - * - = +),
+    fiting perfectly what we want. now a few ilustrations to show better what we
+    are talking about.
+
+    \vspace{3cm}
+
+    \begin{figure}[h]
+        \centering
         \begin{tikzpicture}
             \begin{axis}[
+                         title=Gradient decend movement description,
                          xlabel=$\theta$,
                          ylabel=$J(\theta)$,
-                     ymin=0, ymax=25,
+                         ymin=-5, ymax=25,
                          xtick=\empty,
@@
                         ]
-            \addplot[mark=none, color=red]{x^2};
+                \addplot[mark=none, color=black]{x^2};
 
-            \draw[thick,->] (axis cs: 4,0) -- node[left]{$\theta_1$} (axis cs: 4,16);
+                \addplot[color=red, only marks]
+                    coordinates{
+                        (4,16)
+                        (-4,16)
+                    };
 
@@
                     coordinates{
-                (4,16)
-                (2,4)
-                (1,1)
                         (0,0)
                     };
+
+                \node[above] at (axis cs: 0,1.5) {Minimum};
+
+                \draw[thick,->] (axis cs: 3.5,4) -- node[right]{+} (axis cs: 5.5,20);
+                \draw[thick,->] (axis cs: -3.5,4) -- node[left]{-} (axis cs: -5.5,20);
+
+                \draw[thick,->] (axis cs: -6,-4) -- node[above]{Increment $\theta$} (axis cs: -0.1,-4);
+                \draw[thick,->] (axis cs: 6,-4) -- node[above]{Increment $\theta$} (axis cs: 0.1,-4);
+
             \end{axis}
         \end{tikzpicture}
-
-    This can be done by calculating the direcction and that we want to follow
+    \end{figure}
 
@@
 
-    Now the first thing to do in order to use the gradient desent in our cost
-    function is replace the function $J(\theta)$ in the formula
+    \newpage
+
+    Now to finish the explanation remember the second part of the example, when
+    you start to walk, with a constant step size, down the hill you start
+    making a big advance because of the inclunation, but when you are getting
+    closer to your friends the distance that you move with just one step is
+    smaller because the slope of the hill is lower. This is just what we are
+    trying to show in the last couple of grafics, notice that when the $\theta$
+    is getting closer to the minimum the arrow that represents the movement is
+    smaller, this is always going to happen because the inclunation of the hill
+    is slower so you move less.\\
+
+    Until the moment what we have done is to show the generic version of the
+    gradient desent, what we are goint to do now is to replace the function
+    $J(\theta)$ for our cost function and apply the derivative to the formula
+    in order to get the gradient desend for our particular case.
+
+    Lets start, the first thing to do in order to use the gradient desent in our
+    cost function is replace the function $J(\theta)$ in the formula
 

### 2016-07-08 LinearRegression/Code/OneParam/Octave/OneParam.m

@@ -1,0 +1,74 @@
+#!/usr/bin/octave -q
+
+global theta = 1;
+global alpha = 0.1;
+global err = 0;
+
+function points = getPoints()
+
+    file_desc = fopen("Data.txt", "r");
+
+    tuple_format = [2 Inf];
+    points = fscanf(file_desc,"%d %d", tuple_format);
+
+    fclose(file_desc);
+
+endfunction
+
+function res = h(x)
+
+    global theta;
+    res = theta * x;
+
+endfunction
+
+function cost_function(points)
+
+    global err;
+
+    num_points = size(points);
+    summary = 0;
+
+    for i=1:num_points(2)
+        x = points(1,i);
+        y = points(2,i);
+
+        summary += (h(x) - y)^2;
+    endfor
+
+    err = summary / (2 * num_points(2));
+
+endfunction
+
+function res = cost_function_derivative(points)
+
+    num_points = size(points);
+    summary = 0;
+
+    for i=1:num_points(2)
+        x = points(1,i);
+        y = points(2,i);
+
+        summary += (h(x) - y) * x;
+    endfor
+
+    res = summary / num_points(2);
+
+endfunction
+
+function gradient_decend(points)
+
+    global theta alpha;
+    theta = theta - alpha * cost_function_derivative(points);
+
+endfunction
+
+points = getPoints();
+
+printf("\nalpha = %d\n\n", alpha);
+
+for i=1:3
+    cost_function(points);
+    printf("theta = %d \t error = %d\n", theta, err);
+    gradient_decend(points);
+end

### 2016-07-11 LinearRegression/Code/OneParam/Octave/OneParam.m
printing floats in octave code

@@ -65,10 +65,10 @@ endfunction
 
 points = getPoints();
 
-printf("\nalpha = %d\n\n", alpha);
+printf("\nalpha = %f\n\n", alpha);
 
 for i=1:3
     cost_function(points);
-    printf("theta = %d \t error = %d\n", theta, err);
+    printf("theta = %f \t error = %f\n", theta, err);
     gradient_decend(points);
 end

### 2016-07-11 LinearRegression/Code/OneParam/C++/OneParam.C
adding code in C++

@@ -1,7 +1,70 @@
 #include <iostream>
+#include <fstream>
+#include <sstream>
+#include <math.h>
+#include <stdlib.h>
 
 using namespace std;
 
+float theta = 1;
+float alpha = 0.1;
+
+float x[3];
+float y[3];
+
+float ** getPoints(ifstream& file){
+    string buffer;
+
+    for(int i=0;i<3;i++){
+        getline(file, buffer);
+        istringstream reader(buffer);
+        reader >> x[i] >> y[i];
+    }
+}
+
+float h(float x){
+    return theta * x;
+}
+
+float cost_function(){
+    int num_points = 3;
+    float summary = 0;
+
+    for(int i=0;i<num_points;i++)
+        summary += pow((h(x[i]) - y[i]), 2);
+
+    return summary / (num_points * 2);
+}
+
+float cost_function_derivative(){
+    int num_points = 3;
+    float summary = 0;
+
+    for(int i=0;i<num_points;i++)
+        summary += (h(x[i]) - y[i]) * x[i];
+
+    return summary / num_points;
+}
+
+void gradient_decend(){
+    theta = theta - alpha * cost_function_derivative();
+}
+
 int main(int argc, char *argv[]){
+    ifstream file(argv[1]);
+    int iterations = atoi(argv[2]);
     
+    getPoints(file);
+
+    cout << "alpha = " << alpha << "\n" << endl;
+
+    float err;
+
+    for(int i=0;i<iterations;i++) {
+        err = cost_function();
+        cout << "theta = " << theta << " error = " << err << endl;
+        gradient_decend();
+    }
+
+    return 0;
 }

### 2016-07-11 LinearRegression/Documentation/OneParam/OneParam.tex
Fixing orthographic mistakes

@@
 
-    What the algorithm will do is to draw a line as closser as posible to all
-    the points in order to get the best aproximation posible to the original
+    What the algorithm will do is to draw a line as closser as possible to all
+    the points in order to get the best aproximation possible to the original
     answers\\
@@
     When the right line is founded we can get an aproximatly answers to
-    inputs, that not necesarly where given in first place, allowing to
+    inputs, that not necessarily where given in first place, allowing to
     get more info from data.\\
@@
                       title={Predictions},
-                      xlabel=$h(x)$,
-                      ylabel=$y$,
+                      xlabel=$x$,
+                      ylabel=$h(x)$,
                       xmin=0, xmax=5,
@@
 
-    In order to get this line we must define a function that give as the answers
-    values, the functionwill be the equation of a line
+    In order to get this line we must define a function that give us the answers
+    values, the function will be the equation of a line
     \begin{equation}
@@
     predict. Our function $h(x)$ gave the results $\{5,10,1\}$ and the correct
-    results are $\{6,9,3\}$ and we want to cuantificate the error asocciated to
+    results are $\{6,9,3\}$ and we want to cuantificate the error associated to
     this two answers.
@@
 
-    To achive this gold we have two posible ways to do it, the first one is to
+    To achive this gold we have two possible ways to do it, the first one is to
     add an absolute value to the calculation and the other is to squaring the
@@
 
-    The result of this operation is the set $\{1,1,4\}$. Exelent, now we will
+    The result of this operation is the set $\{1,1,4\}$. Excelent, now we will
     never have a negative number in out resulting set, notice that our
     answers will be much bigger because we are using square of two as our
-    method to eliminate the negative numbers, actually this doesn't metter
-    because we are lookin for a number that describes the error associated to
+    method to eliminate the negative numbers, actually this doesn't matter
+    because we are looking for a number that describes the error associated to
     the prediction that we are computing so the range of the number is not too
@@
     be more concise because the first set is the result of the function $h(x)$,
-    so we are going to define it as $h(x_i)$ and the second can be decribed as
+    so we are going to define it as $h(x_i)$ and the second can be described as
     $y_i$.
 
-    Now we will define the substration of each set as follows
+    Now we will define the subtraction of each set as follows
 
@@
 
-    Applying properties of sum
+    Moving the m aside
 
@@
     This is the cuadratic error and will be used as our cost function in terms
-    of the parameter $\theta$ to expess the error associated to the prediction
+    of the parameter $\theta$ to express the error associated to the prediction
     of the algorithm with respect to the correct output that is given at the
@@
 
-    The next step is define a way to make the error as small as posible,
-    to do this we will use a famus algorithm called gradient decend.
+    The next step is define a way to make the error as small as possible,
+    to do this we will use a famous algorithm called gradient descend.
     the formula of the algorithm is show bellow
@@
     Before explaining the deeper meaning of this derivative we are going to
-    explain in other words what the gradient decend does, imagin that you are
+    explain in other words what the gradient descend does, imagine that you are
     lost in a jungle that is kind of a valley, and you are trying to find your
     frinds that are in the lowest part of the valley, but you dont know exactly
-    in what point of the valley are you, what do you do?\\
+    in what point you are, what do you do?\\
 
@@
             \begin{axis}[
-                         title=What you have to imagin as a valley,
+                         title=What you have to imagine as a valley,
                          ymin=0, ymax=25,
@@
     example), you should notice that you have a big advance with each step
-    becouse of the high inclination, but when you are aproaching your friends
+    because of the high inclination, but when you are aproaching your friends
     the inclunation is lower and you start to get smaller steps until you get
@@
 
-    Now that we have seen the example lets analyze the formula, the fist symbol
+    Now that we have seen the example lets analyze the formula, the first symbol
     is the $\theta$ that in the example is your position but in the
@@
     which means that we are updating the value or assigning a new one,
-    this make sence because we are trying to get closer to our friends which
+    this make sence because we are trying to get closer to our friends, which
     means update the cost function ($J(\theta)$) to the minimum. Then we
@@
     Remember that the first thing we do in the example was looked at the
-    inclunation of the hill where we are, this in order to know if we should
-    move right or left. Now, as you may know a derivative is basicaly the
+    inclination of the hill where we are, this in order to know if we should
+    move right or left. Now, as you may know a derivative is basically the
     slope of a function in a point, this is perfect for us because fit our
-    request perfectly. Lets imaging that we start at the right side of the
+    request perfectly. Lets imagine that we start at the right side of the
     valley, so we what to go left, now mathematically the derivative of the
@@
     substracted to the actual value of $\theta$, just as we want.
-    Now lets imaging that we end up in the left side of the valley, in that case
+    Now lets imagine that we end up in the left side of the valley, in that case
     we want to go right, now mathematically the derivative of a point that is
@@
     and then added to the actual value of $\theta$ (remember that - * - = +),
-    fiting perfectly what we want. now a few ilustrations to show better what we
+    fitting perfectly what we want. now a few ilustrations to show better what we
     are talking about.
@@
             \begin{axis}[
-                         title=Gradient decend movement description,
+                         title=Gradient descend movement description,
                          xlabel=$\theta$,
@@
         \begin{axis}[
-                     title={Gradient decend with positive derivative},
+                     title={Gradient descend with positive derivative},
                      xlabel=$\theta$,
@@
         \begin{axis}[
-                     title={Gradient decend with negative derivative},
+                     title={Gradient descend with negative derivative},
                      xlabel=$\theta$,
@@
     is getting closer to the minimum the arrow that represents the movement is
-    smaller, this is always going to happen because the inclunation of the hill
+    smaller, this is always going to happen because the inclination of the hill
     is slower so you move less.\\
@@
     Now the factor of the denominator has no influence on the function so it
-    can be tranlated outside of the derivative
+    can be translated outside of the derivative
 
@@
     The derivative of the summatory is equal to the derivative of each term and
-    then add each result, so we can insert the dirivative into the summary
+    then add each result, so we can insert the derivative into the summary
 
@@
     (the 2 that we added at the begining), this is with the objective
-    to cancel the 2 in the numerator that result of the derivative. Finaly
+    to cancel the 2 in the numerator that result of the derivative. Finally
     this is the resulting function that we are going to use to minimize the

### 2016-07-16 OctaveExamples/Array/Array.m
Removing Octave examples

@@ -1,5 +1,0 @@
-A = [1,2,3,4]
-A(1)
-
-B = [1,2;3,4]
-B(1,2)

### 2016-07-16 OctaveExamples/Conditional/Conditional.m

@@ -1,7 +1,0 @@
-A=[1,2,3]
-
-if (A(1) == 1)
-    disp("true")
-else
-    disp("false")
-endif

### 2016-07-16 OctaveExamples/Functions/Function.m

@@ -1,5 +1,0 @@
-function first()
-    disp("This is the first function")
-endfunction
-
-first()

### 2016-07-16 OctaveExamples/Global/Global.m

@@ -1,8 +1,0 @@
-global ar = 1;
-
-function ret = test()
-    global ar;
-    ret = ar + 1;
-endfunction
-
-test()

### 2016-07-16 OctaveExamples/Loop/For.m

@@ -1,3 +1,0 @@
-for i = 1:10
-    disp(i)
-endfor

### 2016-07-16 OctaveExamples/Loop/While.m

@@ -1,5 +1,0 @@
-i = 0
-
-while (i<=10)
-    i++
-endwhile

### 2016-07-16 OctaveExamples/Print/Print.m

@@ -1,1 +1,0 @@
-disp("hola")

### 2016-07-20 LinearRegression/Code/TwoParam/C++/TwoParam.C
adding first implementation of linear regression with two parameters

@@ -1,7 +1,90 @@
 #include <iostream>
+#include <fstream>
+#include <sstream>
+#include <math.h>
+#include <stdlib.h>
 
 using namespace std;
 
+float theta_0 = 1;
+float theta_1 = 1;
+float alpha = 0.1;
+
+float x[3];
+float y[3];
+
+float ** getPoints(ifstream& file){
+    string buffer;
+
+    for(int i=0;i<3;i++){
+        getline(file, buffer);
+        istringstream reader(buffer);
+        reader >> x[i] >> y[i];
+    }
+
+}
+
+float h(float x){
+    return theta_0 + theta_1 * x;
+}
+
+float cost_function(){
+    int num_points = 3;
+    float summary = 0;
+
+    for(int i=0;i<num_points;i++)
+        summary += pow((h(x[i]) - y[i]), 2);
+
+    return summary / (num_points * 2);
+}
+
+float cost_function_derivative_theta0(){
+    int num_points = 3;
+    float summary = 0;
+
+    for(int i=0;i<num_points;i++)
+        summary += (h(x[i]) - y[i]) * x[i];
+
+    return summary / num_points;
+}
+
+float cost_function_derivative_theta1(){
+    int num_points = 3;
+    float summary = 0;
+
+    for(int i=0;i<num_points;i++)
+        summary += (h(x[i]) - y[i]);
+
+    return summary / num_points;
+}
+
+void gradient_decend(){
+    float temp_0 = theta_0 - alpha * cost_function_derivative_theta0();
+    float temp_1 = theta_1 - alpha * cost_function_derivative_theta1();
+
+    theta_0 = temp_0;
+    theta_1 = temp_1;
+}
+
 int main(int argc, char *argv[]){
-    
+    ifstream file(argv[1]);
+    int iterations = atoi(argv[2]);
+
+    getPoints(file);
+
+    cout << "alpha = " << alpha << "\n" << endl;
+
+    float err;
+
+    for(int i=0;i<iterations;i++) {
+        err = cost_function();
+        cout << "theta_0 = " << theta_0 << endl;
+        cout << "theta_1 = " << theta_1 << endl;
+        cout << "error = " << err << endl;
+        cout << endl;
+        gradient_decend();
+    }
+
+    return 0;
+
 }

### 2016-07-20 LinearRegression/Code/TwoParam/Octave/TwoParam.m
adding first implementation of linear regression with two parameters in Octave

@@ -1,1 +1,101 @@
 #!/usr/bin/octave -q
+
+global theta_0 = 1;
+global theta_1 = 1;
+global alpha = 0.1;
+global err = 0;
+
+function points = getPoints()
+
+    file_desc = fopen("Data.txt", "r");
+
+    tuple_format = [2 Inf];
+    points = fscanf(file_desc,"%d %d", tuple_format);
+
+    fclose(file_desc);
+
+endfunction
+
+function res = h(x)
+
+    global theta_0 theta_1;
+
+    res = theta_0 + theta_1 * x;
+
+endfunction
+
+function cost_function(points)
+
+    global err;
+
+    num_points = size(points);
+    summary = 0;
+
+    for i=1:num_points(2)
+        x = points(1,i);
+        y = points(2,i);
+
+        summary += (h(x) - y)^2;
+    endfor
+
+    err = summary / (2 * num_points(2));
+
+endfunction
+
+function res = cost_function_derivative_theta_0(points)
+
+    num_points = size(points);
+    summary = 0;
+
+    for i=1:num_points(2)
+        x = points(1,i);
+        y = points(2,i);
+
+        summary += (h(x) - y);
+    endfor
+
+    res = summary / num_points(2);
+
+endfunction
+
+function res = cost_function_derivative_theta_1(points)
+
+    num_points = size(points);
+    summary = 0;
+
+    for i=1:num_points(2)
+        x = points(1,i);
+        y = points(2,i);
+
+        summary += (h(x) - y) * x;
+    endfor
+
+    res = summary / num_points(2);
+
+endfunction
+
+function gradient_decend(points)
+
+    global theta_0 theta_1 alpha;
+    temp_0 = theta_0 - alpha * cost_function_derivative_theta_0(points);
+    temp_1 = theta_1 - alpha * cost_function_derivative_theta_1(points);
+
+    theta_0 = temp_0;
+    theta_1 = temp_1;
+
+endfunction
+
+points = getPoints();
+
+printf("\nalpha = %f\n\n", alpha);
+
+for i=1:3
+    cost_function(points);
+
+    printf("theta_0 = %f\n", theta_0);
+    printf("theta_1 = %f\n", theta_1);
+    printf("error = %f\n", err);
+    printf("\n");
+
+    gradient_decend(points);
+end

### 2018-03-02 Dnn/Main.cpp
First version of Dnn complete

@@ -7,7 +7,7 @@ using namespace std;
 int main(int argc, char *argv[]){
 
     int num_layers = 3;
-    int structure[num_layers] = {2,2,2}; //first layer are inputs
+    int layers[num_layers] = {2,2,2}; //first layer are inputs
 
     double input[3] = {0.05, 0.10, 1}; //two inputs plus bias
 
@@ -19,7 +19,7 @@ int main(int argc, char *argv[]){
 
     double labels[2] = {0.01, 0.99};
 
-    NNet * nnet = new NNet(structure, num_layers);
+    NNet * nnet = new NNet(layers, num_layers);
 
     nnet->set_input(input);
     nnet->set_weights(weights_1, 0);

### 2018-03-02 Dnn/NNet.cpp

@@ -1,6 +1,10 @@
 #include <math.h>
 #include "NNet.h"
 
+#include <iomanip>
+#include <iostream>
+using namespace std;
+
 NNet::NNet(int * layers, int num_layers){
     this->layers = layers;
     this->num_layers = num_layers;
@@ -12,6 +16,8 @@ NNet::NNet(int * layers, int num_layers){
     f = new double*[num_layers]; //will contain the value of all neurons
 
     functions = new Function*[num_layers+1]; //will contain activations and loss function
+                                             //the first layer are inputs without
+                                             //activation, so the first possition is null
 }
 
 NNet::~NNet(){
@@ -87,39 +93,47 @@ double NNet::loss(double * y_hat){
 
 void NNet::backward(){
 
-    double dE_dy, dy_do, dE_do;
-    double do_dz, do_dw, dE_dw;
+    double dE_dy, dy_do, dE_do, do_dz, do_dw, dE_dw;
     double* w;
 
-    int num_neur_actual, num_neur_next;
+    int num_neur_actual, num_neur_next, num_neur_past;
+    int con_layer, bias_pos;
 
     for(int layer=last_layer;layer>0;layer--){ //for each layer
 
         num_neur_actual = layers[layer]; //height of the matrix
         num_neur_next = layers[layer-1]+1; //width of the matrix
-        b[layer-1] = new double[num_neur_actual]; //allocate space for backward information
+
+        con_layer = layer-1; //index for W and b
+        bias_pos = num_neur_next-1; //bias possition in the matrix of weights
+
+        b[con_layer] = new double[num_neur_actual * num_neur_next]; //allocate space for backward information
 
         for(int i=0;i<num_neur_actual;i++){ //for each neuron
 
             if(layer == last_layer) dE_dy = functions[last_function]->derivative(f[layer], i); //deriv error respect to final output
             else{
-                dE_dy = 0;
-                for(int j=0;j<num_neur_actual-1;j++){
-                    dE_do = b[layer][i*num_neur_next + 2];
-                    do_dz = W[layer][j*num_neur_next + i];
-                    dE_dy += dE_do * do_dz;
+                dE_dy = 0; //reset the deriv error respect to the output
+                num_neur_past = layers[layer+1]; //number of forward paths, bias is not a forward pass
+
+                for(int j=0;j<num_neur_past;j++){ //for each forward path
[…]

### 2018-10-29 DimReduction/Mds/Mds.m
Adding first part of Mds code

@@ -1,0 +1,23 @@
+% colordef none;
+% 
+% n = 1000;
+% [x y z] = build_swiss_roll(n);
+% data = [x y z];
+% 
+% colors = get_cuadrant_colors(x,z);
+% plot_swiss_roll(x,y,z, colors, [-15 15]);
+
+n = 10;
+p = 2;
+
+H = eye(n) - ones(n)/n;
+A = -0.5 * D.^2;
+B = H * A * H;
+
+[V, L] = eigs(B,p+3,'LR')
+L_half = sqrt(L);
+Y_full = V * L_half;
+Y = Y(:,1:p)
+
+plot(Y(:,1), Y(:,2), '.')
+text(Y(:,1), Y(:,2), names)

### 2018-10-30 DimReduction/Lle/Documentation/Article_lle.tex
Adding new version of dimensionality reduction

\documentclass[12pt,journal]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\title{A weighted data representation based approach for manifold learning}
\author{A. Salgado - O. Lucía Quintero M.}
\maketitle

\begin{abstract}
    We implemented the weighted data representation based in KNN and minimization
    of the euclidean distance between representation. Finally the local embedding
    is tested on 3-dimensional manifolds.
\end{abstract}

\section{Introduction}

\section{Theorical framework}

\subsection{First optimization problem}

    \begin{equation*}
        \begin{aligned}
            \underset{w_{ij}}{\text{min}} \quad \sum_{i=1}^n & \lVert x_i - \sum_{j=1}^k w_{ij} N(x_i)_j \rVert^2 \\
            & \sum_{j=1}^k w_{ij} = 1
        \end{aligned}
    \end{equation*}

    \subsection{Case for an arbitrary $i$}

    define

    \[
        V_i = 
        \begin{bmatrix}
            |      &  |     &       & | \\
            N(x_i)_1 & N(x_i)_2 & \dots & N(x_i)_k \\
            |      &  |     &       & |
        \end{bmatrix}_{d \times k}
    \]
    \[
        W_i =
        \begin{bmatrix}
            w_{i1}\\
            w_{i2}\\
            \dots \\
            w_{ik}\\
        \end{bmatrix}_{k \times 1}
        \hspace{1cm}
        e = 
        \begin{bmatrix} 
        1\\
        1\\
        \dots\\
        1
        \end{bmatrix}_{k \times 1}
    \]

    then

    \begin{equation*}
        \sum_{j=1}^k w_{ij} N(x_i)_j = V_i W_i
    \end{equation*}

    \begin{equation*}
            \underset{w_{ij}}{\text{min}} \lVert x_i - \sum_{j=1}^k w_{ij} N(x_i)_j \rVert^2
            = 
            \underset{W_i}{\text{min}} \lVert x_i - V_i W_i \rVert^2
    \end{equation*}

    define

    \[
        x_i e^t =
        \begin{bmatrix}
            x_1 \\
            x_2 \\
            \dots \\
            x_d \\
        \end{bmatrix}_{d \times 1}
        \begin{bmatrix}
            1 & \dots & 1
        \end{bmatrix}_{1 \times k}
        =
        \begin{bmatrix}
            |   &       & | \\
            x_i & \dots & x_i\\
            |   &       & |
        \end{bmatrix}_{d \times k}
    \]\\

    Now $x_i$ can be defined as

    \begin{equation*}
        x_i = x_i e^t W_i
    \end{equation*}

    Hence

    \begin{equation*}
        \begin{aligned}
            \underset{W_i}{\text{min}} \lVert x_i - V_i W_i \rVert^2
            &=
            \underset{W_i}{\text{min}} \quad \lVert x_i e^t W_i - V_i W_i \rVert^2\\
            &=
            \underset{W_i}{\text{min}} \quad \lVert (x_i e^t - V_i) W_i \rVert^2\\
            &=
            \underset{W_i}{\text{min}} \quad [(x_i e^t - V_i) W_i]^t [(x_i e^t - V_i) W_i]\\
            &=
            \underset{W_i}{\text{min}} \quad W_i^t(x_i e^t - V_i)^t (x_i e^t - V_i) W_i\\
        \end{aligned}
    \end{equation*}

    Define a matrix $G$ as

    \begin{equation*}
        G = (x_i e^t - V_i)^t (x_i e^t - V_i)
    \end{equation*}

    Redefine the problem as

    \begin{equation*}
            \underset{W_i}{\text{min}} \quad W_i^t(x_i e^t - V_i)^t (x_i e^t - V_i) W_i
            =
            \underset{W_i}{\text{min}} \quad W_i^t G W_i
    \end{equation*}

    To solve this problem the Lagrange multipliers are used. First redefine the
    restriction as

    \begin{equation*}
        \sum_{j=1}^k w_{ij} = 1 = e^t W_i
    \end{equation*}

    Now

    \begin{equation*}
        \begin{aligned}
            \mathcal{L}(W_i, \lambda) &= W_i^t G W_i - \lambda (e^t W_i - 1)\\
            \frac{\delta \mathcal{L}}{\delta W_i} &= 2 G W_i - \lambda e
        \end{aligned}
    \end{equation*}

    \begin{equation*}
        \begin{aligned}
            2 G W_i - \lambda e &= 0 \\
            2 G W_i &= \lambda e \\
            G W_i &= \frac{1}{2} \lambda e \\
            G^{-1} G W_i &= G^{-1} \frac{1}{2} \lambda e \\
            W_i &= \frac{1}{2} \lambda G^{-1} e \\
        \end{aligned}
    \end{equation*}

    Now if $\lambda$ is wrongly chosen, then the result would be a scaled version
    of the true value. Then, $\lambda$ can be chosen arbitrarily, and after $W_i$
    is computed a rescale procedure can be conducted in order to satisfy the
    restriction.\\

    For simplicity choose $\lambda = 2$, then

    \begin{equation*}
        \begin{aligned}
            W_i &= \frac{1}{2} \lambda G^{-1} e = G^{-1} e \\
            G W_i &= e \\
        \end{aligned}
    \end{equation*}

    So the problem is reduced to solve a linear system of equations. Now if
    after build the matrix G, it turns out that is a singular matrix the
    following computation can be used to make it invertible

    \begin{equation*}
            G = G + \sigma I
    \end{equation*}

    where sigma is a small value.

    \subsection{Second optimization problem}

    \begin{equation*}
        \begin{aligned}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n & \lVert y_i - \sum_{j=1}^n w_{ij} y_j \rVert^2 \\
            & YY^t = I
        \end{aligned}
    \end{equation*}

    Where

    \[
        Y = 
        \begin{bmatrix}
            |   &  |  &       & |   \\
            y_1 & y_2 & \dots & y_n \\
            |   &  |  &       & |
        \end{bmatrix}_{p \times n}
        \hspace{0.5cm}
    \]

    Now, define

    \[
        W = 
        \begin{bmatrix}
            |   &  |  &       & |   \\
            w_1 & w_2 & \dots & w_n \\
            |   &  |  &       & |
        \end{bmatrix}_{n \times n}
        where
        \hspace{0.5cm}
        W_i = 
        \begin{bmatrix}
            0 \\
            \vdots \\ 
            w_1 \\
            \vdots \\ 
            w_k \\
            \vdots \\ 
            0 \\
        \end{bmatrix}_{n \times 1}
    \]

    Where $W_i$ is a vector that contains the value of the weight that correspond
    to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
    else. This means that

    \begin{equation*}
        X W_i = \sum_{j=1}^k w_{ij} x_j
    \end{equation*}

    Rewrite the problem as

    \begin{equation*}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - \sum_{j=1}^n w_{ij} y_j \rVert^2
            =
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - Y W_i \rVert^2
    \end{equation*}

    Also define

    \[
        I_i = 
        \begin{bmatrix}
            0 \\
            \vdots \\ 
            1 \\
            \vdots \\ 
            0 \\
        \end{bmatrix}_{n \times n}
        \hspace{0.5cm}
    \]

    Which is a vector that has a $1$ in the $i$ position and $0$ anywhere else.
    With this new vector we can redefine the problem as

    \begin{equation*}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - Y W_i \rVert^2
            =
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert YI_i - Y W_i \rVert^2
    \end{equation*}

    And now, since this expression is adding the norm of each vector in the
    matrix, the optimization problem can be rewritten to minimize the norm of
    the entire matrix

    \begin{equation*}
            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert YI_i - Y W_i \rVert^2
            =
            \underset{Y}{\text{min}} \quad \lVert YI - Y W \rVert^2
    \end{equation*}

    So, this optimization problem can be solved as follows

    \begin{equation*}
        \begin{aligned}
            \lVert Y I - Y W \rVert^2 &= \lVert Y (I - W) \rVert^2\\
            &= Tr[(Y [I-W])^t (Y [I-W])]\\
            &= Tr[(I-W)^t Y^t Y (I-W)]\\
            &= Tr[(I-W)^t Y^t Y (I-W)]\\
            &= Tr[(I-W) (I-W)^t Y^t Y]\\
            &= Tr[Y (I-W) (I-W)^t Y^t]\\
        \end{aligned}
    \end{equation*}

    At this point define a matrix $M$ as

    \begin{equation*}
        M = (I-W) (I-W)^t
    \end{equation*}

    Redefine the optimization problem as

    \begin{equation*}
        \begin{aligned}
            \underset{Y}{\text{min}} \quad \lVert YI - Y W \rVert^2
            &= 
            \underset{Y}{\text{min}} \quad Tr(Y M Y^t)\\
            &=
            \underset{Y}{\text{min}} \quad Tr(M Y^t Y)
        \end{aligned}
    \end{equation*}

    To solve this problem the Lagrange multipliers are used.

    \begin{equation*}
        \begin{aligned}
            \mathcal{L}(Y, \lambda) &= Tr( M Y^t Y) - \lambda (Y Y^t - I)\\
            \frac{\delta \mathcal{L}}{\delta W_i} &= Y M^t + Y M - 2 \lambda Y
        \end{aligned}
    \end{equation*}

    We know that matrix $M$ is symmetric

    \begin{equation*}
        \begin{aligned}
            M^t &= [(I-W) (I-W)^t]^t\\
                &= [(I-W)^t]^t (I-W)^t\\
                &= (I-W) (I-W)^t\\
                &= M
        \end{aligned}
    \end{equation*}

    Hence

    \begin{equation*}
        \begin{aligned}
            \frac{\delta \mathcal{L}}{\delta W_i} &= Y M^t + Y M - 2 \lambda Y\\
                                                  &= Y M + Y M - 2 \lambda Y\\
                                                  &= 2 Y M - 2 \lambda Y \\
                              2 Y M - 2 \lambda Y &= 0 \\
                                            2 Y M &= 2 \lambda Y \\
                                              Y M &= \lambda Y \\
                                              (Y M)^t &= (\lambda Y)^t \\
                                              M^t Y^t &= Y^t \lambda \\
                                              M Y^t &= \lambda Y^t \\
        \end{aligned}
    \end{equation*}

\section{Algorithm implementation}

\section{Results}

\section{Conclusions}

\end{document}

### 2018-10-30 DimReduction/Mds/Documentation/Article_mds.tex
Adding new version of dimensionality reduction

\documentclass[12pt,journal]{IEEEtran}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}

\title{A theorical-practical approach to manifold learning}
\author{A. Salgado - O. Lucía Quintero M.}
\maketitle

\begin{abstract}
    This work presents an theorical approximation to the problem and provide an
    algorithmic solution easy and understandable. The approach to the problem is
    reconsider by the construction of the matrix $D^x$ with KNN. Creating the
    possibility to improve the results using other ways to calculate the distance
    matrix
\end{abstract}

\section{Introduction}

\section{Theorical framework}

    \subsection{Points from distances}

    The set up of the problem is a matrix of euclidean distances $D$, where each
    entry is the square of the distance between two points in a $d$ dimensional
    space.

    \[
        D =
        \begin{bmatrix}
            d_{11}^2 & \cdots & d_{1n}^2\\
            \vdots   & \ddots & \vdots  \\
            d_{n1}^2 & \ddots & d_{nn}^2\\
        \end{bmatrix}_{n x n}
    \]

    where

    \begin{equation*}
        d_{ij}^2 = (x_{i1}-x_{j1})^2 - (x_{i2}-x_{j2})^2 \cdots (x_{id}-x_{jd})^2
    \end{equation*}

    The unknown of the problem is a matrix $X$ that cotains the position of the
    $n$ points in the $d$ dimensional space.

    \[
        X =
        \begin{bmatrix}
            x_{11} & \cdots & x_{1n}\\
            \vdots & \ddots & \vdots\\
            x_{d1} & \ddots & x_{dn}\\
        \end{bmatrix}_{d x n}
    \]

    But there is a restriction to make sure that the matrix is centered at 0.

    \begin{equation*} 
        \sum_{i=1}^n x_{is} = 0 \quad with \quad s = 1, \cdots, d
    \end{equation*}

    \subsection{Solution}

    First rewrite each entry of matrix $D$ in a vector form

    \begin{equation*}
        d_{ij}^2 = (x_i-x_j)^t (x_i-x_j)
    \end{equation*}

    Then expand distance $d_{ij}^2$

    \begin{equation*}
        \begin{aligned}
            (x_i-x_j)^t (x_i-x_j) &= (x_i^t-x_j^t) (x_i-x_j)\\
                                  &= x_i^t x_i - x_i^t x_j - x_j^t x_i + x_j^t x_j\\
                         d_{ij}^2 &= x_i^t x_i - 2 x_i^t x_j + x_j^t x_j\\
        \end{aligned}
    \end{equation*}

    The goal is to express all the dot products of $x$ vectors in terms of the
    $d_{ij}^2$ distances. Consider the following expression

    \begin{equation*}
        \begin{aligned}
            \frac{1}{n} \sum_{i=1}^n d_{ij}^2 &= \frac{1}{n} \sum_{i=1}^n [x_i^t x_i - 2 x_i^t x_j + x_j^t x_j]\\
            &=  \frac{1}{n} \sum_{i=1}^n x_i^t x_i - \frac{2}{n} \sum_{i=1}^n x_i^t x_j + \frac{1}{n} \sum_{i=1}^n x_j^t x_j\\
            &=  \frac{1}{n} \sum_{i=1}^n x_i^t x_i - \frac{2}{n} \sum_{i=1}^n x_i^t x_j + \frac{n}{n} x_j^t x_j \\
        \end{aligned}
    \end{equation*}

    Remember that

    \begin{equation*}
        \begin{aligned}
              \sum_{i=1}^n x_{is} &= 0 \quad with \quad s = 1, \cdots, d\\
            \sum_{i=1}^n x_i^t x_j &= \sum_{i=1}^n \sum_{r=1}^d x_{ir} x_{jr}\\
                                   &= \sum_{r=1}^d \sum_{i=1}^n x_{ir} x_{jr}\\
                                   &= \sum_{r=1}^d x_{jr} \sum_{i=1}^n x_{ir}\\
                                   &= \sum_{r=1}^d x_{jr} 0\\
                                   &= 0\\
        \end{aligned}
    \end{equation*}

    Hence

    \begin{equation*}
        \begin{aligned}
            \frac{1}{n} \sum_{i=1}^n d_{ij}^2
            &= 
            \frac{1}{n} \sum_{i=1}^n x_i^t x_i - \frac{2(0)}{n} + x_j^t x_j\\
            &= 
            \frac{1}{n} \sum_{i=1}^n x_i^t x_i + x_j^t x_j\\
            x_j^t x_j &= \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - \frac{1}{n} \sum_{i=1}^n x_i^t x_i
        \end{aligned}
    \end{equation*}

    Now in the other hand, the sum over $j$ would be

    \begin{equation*}
        \begin{aligned}
            \frac{1}{n} \sum_{j=1}^n d_{ij}^2
            &=
            \frac{1}{n} \sum_{j=1}^n [x_i^t x_i - 2 x_i^t x_j + x_j^t x_j]\\
            &= 
            x_i^t x_i + \frac{1}{n} \sum_{j=1}^n x_j^t x_j\\
            x_i^t x_i &= \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - \frac{1}{n} \sum_{j=1}^n x_j^t x_j\\
        \end{aligned}
    \end{equation*}

    Finally, the first approximation to the problem can be constructed

    \begin{equation*}
        \begin{aligned}
            d_{ij}^2 &= x_i^t x_i - 2 x_i^t x_j + x_j^t x_j\\
            2 x_i^t x_j &= x_i^t x_i - d_{ij}^2 + x_j^t x_j\\
            x_i^t x_j &= \frac{1}{2} (x_i^t x_i - d_{ij}^2 + x_j^t x_j)\\
            \text{where}\\
            x_i^t x_i &= \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - \frac{1}{n} \sum_{j=1}^n x_j^t x_j\\
            x_j^t x_j &= \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - \frac{1}{n} \sum_{i=1}^n x_i^t x_i
        \end{aligned}
    \end{equation*}

    Notice that

    \begin{equation*}
        \sum_{i=1}^n x_i^t x_i = \sum_{j=1}^n x_j^t x_j
    \end{equation*}

    So

    \begin{equation*}
        \begin{aligned}
            x_i^t x_j&=
            \frac{1}{2} \left[ \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - d_{ij}^2 + \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - \frac{2}{n} \sum_{i=1}^n x_i^t x_i \right]\\
        \end{aligned}
    \end{equation*}

    This is almost expressed in terms of the distances, the only problem is the
    last term. To express that part consider the following expression

    \begin{equation*}
        \begin{aligned}
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n d_{ij}^2
            &= 
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n [x_i^t x_i - 2 x_i^t x_j + x_j^t x_j]\\
            &= 
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i - \frac{2(0)}{n^2} +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
            &= 
            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
            &= 
            \frac{n}{n^2}\sum_{i=1}^n x_i^t x_i +  \frac{n}{n^2}\sum_{j=1}^n x_j^t x_j\\
            &= 
            \frac{1}{n}\sum_{i=1}^n x_i^t x_i +  \frac{1}{n}\sum_{j=1}^n x_j^t x_j\\
            &= 
            \frac{2}{n}\sum_{i=1}^n x_i^t x_i\\
        \end{aligned}
    \end{equation*}

    Now the expression can be completed
    \begin{equation*}
        \begin{aligned}
            x_i^t x_j
            &=
            \frac{1}{2} \left[ \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - d_{ij}^2 + \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n d_{ij}^2 \right]\\
        \end{aligned}
    \end{equation*}

    This computation can be expressed in matrix form to simplify the formula.
    First define

    \[
        b_{ij} = x_i^t x_j
        \hspace{1cm}
        e =
        \begin{bmatrix}
            1\\
            \vdots\\
            1
        \end{bmatrix}_{nx1}
        \hspace{1cm}
        a_{ij} = -\frac{1}{2} d_{ij}^2
    \]

    Now a matrix $B$ can be defined as

    \begin{equation*}
        B = H A H \quad \text{where} \quad H = I - \frac{1}{n} e e^t
    \end{equation*}

    This formula gives a way to compute all the dot products between the points
    that makes up the matrix $X$. Because of this the matrix $B$ can also
    be expressed as

    \begin{equation*}
        B = X^tX
    \end{equation*}

    Now, as shown in section \ref{orthogonal}, $B$ can be decomposed using the
    spectral decomposition. And by doing that the problem can be solved as follows

    \begin{equation*}
        \begin{aligned}
            B &= V \Sigma V^t\\
              &= V \Sigma^{1/2} \Sigma^{1/2} V^t\\
              &= V \Sigma^{1/2} \Sigma^{1/2} V^t\\
              &= X^t X\\
            X & = \Sigma^{1/2} V^t
        \end{aligned}
    \end{equation*}

    Which gives a way to compute all the points of the matrix $X$.

    \section{Mds optimization problem}

    \begin{equation*}
        \begin{aligned}
            \underset{y}{\text{min}}  & \sum_{i=1}^n \sum_{j=1}^n ( d^{(x)}_{ij} - d^{(x)}_{ij} )^2 \\
            & d_{ij}^{(x)} = \lVert x_i - x_j \rVert\\
            & d_{ij}^{(y)} = \lVert y_i - y_j \rVert
        \end{aligned}
    \end{equation*}\\

    if $d_{ij}^{(x)}$ es euclidean, then $d_{ij}^{(x)} = x_i^t x_j$, and hence

    \begin{equation*}
        \underset{y}{\text{min}} \sum_{i=1}^n \sum_{j=1}^n ( x_i^t x_j - y_i^t y_j )^2 \\
    \end{equation*}

    or in matrix form

    \begin{equation*}
        \underset{Y}{\text{min}} \quad \lVert X^t X - Y^t Y \rVert^2 \\
    \end{equation*}

    As shown in section \ref{norm_trace}, the norm of a matrix can be transformed
    to a trace, so the problem can be redefined as

    \begin{equation*}
        \begin{aligned}
            A = X^tX &- Y^tY\\
            \underset{Y}{\text{min}} \quad \lVert X^t X - Y^t Y \rVert^2
            &=
            \underset{Y}{\text{min}} \quad \lVert A \rVert^2\\
            &=
            \underset{Y}{\text{min}} \quad Tr(A^tA) \\
        \end{aligned}
    \end{equation*}\\

    Now

    \begin{equation*}
        \begin{aligned}
            A^t &= (X^tX - Y^tY)^t\\
                &= (X^tX)^t - (Y^tY)^t\\
                &= X^tX - Y^tY\\
                &= A
        \end{aligned}
    \end{equation*}

    Which means that $A^tA = AA = A^2$, and hence

    \begin{equation*}
        \begin{aligned}
            \underset{Y}{\text{min}} \quad Tr(A^tA)
            &=
            \underset{Y}{\text{min}} \quad Tr(A^2)\\
            &=
            \underset{Y}{\text{min}} \quad Tr[(X^tX - Y^tY)^2]
        \end{aligned}
    \end{equation*}\\

    now as shown in section \ref{spectral_decomp}, $X^tX$ and $Y^tY$ can be
    expresed as

    \begin{equation*}
        X^tX = V\Lambda V^t \hspace{1cm} Y^tY = Q \hat{\Lambda} Q^t
    \end{equation*}

    Then

    \begin{equation*}
        \underset{Y}{\text{min}} \quad Tr[(X^tX - Y^tY)^2]
        =
        \underset{Q,\hat{\Lambda}}{\text{min}} \quad Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]
    \end{equation*}

    We know that $V^tV = I = VV^t$, then using the proof presented in section \ref{circular_trace}

    \begin{equation*}
        \begin{aligned}
            A =& V\Lambda V^t - Q \hat{\Lambda} Q^t\\
            Tr[(V^tAV)^2] &= Tr[V^tAVV^tAV]\\
                          &= Tr[V^tAIAV]\\
                          &= Tr[AIAVV^t]\\
                          &= Tr[AIAI]\\
                          &= Tr[AA]\\
                          &= Tr[A^2]\\
        \end{aligned}
    \end{equation*}\\

    Which implies that

    \begin{equation*}
        \begin{aligned}
            Tr[(V^tAV)^2] &= Tr[A^2]\\
            Tr[(V^t(V\Lambda V^t - Q \hat{\Lambda} Q^t)V)^2] &= Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]\\
            Tr[(V^tV\Lambda V^tV - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
            Tr[(I\Lambda I - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
        \end{aligned}
    \end{equation*}

    Now

    \begin{equation*}
        \begin{aligned}
            G &= V^tQ\\
            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &= Tr[(\Lambda - G \hat{\Lambda} G^t)^2]\\
        \end{aligned}
    \end{equation*}\\

    Which means that the problem can be redefined one more time

    \begin{equation*}
        \begin{aligned}
        \underset{Q,\hat{\Lambda}}{\text{min}} \quad Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]
        &=
        \underset{G,\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
        \end{aligned}
    \end{equation*}\\

    Now

    \begin{equation*}
        \begin{aligned}
        Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
        &=
        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + (G \hat{\Lambda} G^t)^2]\\
        &=
        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} G^t G \hat{\Lambda} G^t]\\
        \end{aligned}
    \end{equation*}

    But (Knowing that $Q^tQ = I = QQ^t$)

    \begin{equation*}
        \begin{aligned}
            G^tG &= (V^tQ)^t V^tQ\\
                 &= (Q^tV) V^tQ\\
                 &= Q^tQ\\
                 &= I\\ \\
            GG^t &= V^tQ (V^tQ)^t\\
                 &= V^tQ (Q^tV)\\
                 &= V^tV\\
                 &= I\\ \\
        \end{aligned}
    \end{equation*}

    Hence

    \begin{equation*}
        \begin{aligned}
        Tr[(\Lambda - G \hat{\Lambda} G^t G)^2 ]
        &=
        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} \hat{\Lambda} G^t]\\
        =
        Tr[&\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda}^2 G^t]\\
        =
        Tr[&\Lambda^2]- 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[G \hat{\Lambda}^2 G^t]\\
        =
        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2 G^tG]\\
        =
        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2]\\
        =
        Tr[&\Lambda^2] + Tr[ \hat{\Lambda}^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t]\\
        \end{aligned}
    \end{equation*}

    \subsection{Dark part}

    We know that $\frac{\delta Tr[\Lambda G \hat{\Lambda} G^t]}{\delta G} = 2\Lambda G \hat{\Lambda}$\\\\
    We also know that $\frac{\delta Tr(\Lambda^2)}{\delta G} = 0 = \frac{\delta Tr(\hat{\Lambda}^2)}{\delta G}$\\\\
    Finally we know that the answer for a fixed $\hat{\Lambda}$ is $G=I$\\ \\

    Which implies that

    \begin{equation*}
        \begin{aligned}
        G = I = V^tQ\\
        V^tQ = I\\
        V(V^tQ) = VI\\
        Q = V
        \end{aligned}
    \end{equation*}

    Now, coming back to the original problem

    \begin{equation*}
        \begin{aligned}
            \underset{G,\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
            &=
            \underset{\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - \hat{\Lambda})^2]
        \end{aligned}
    \end{equation*}

    So we can take the first $p$ eigenvalues of $\Lambda$ to define $\hat{\Lambda}$,
    and finally

    \begin{equation*}
        \begin{aligned}
            Y^tY &= Q \hat{\Lambda} Q^t\\
            &=
            V \hat{\Lambda} V^t\\
            &=
            V \hat{\Lambda}^{1/2} \hat{\Lambda}^{1/2} V^t\\
        \end{aligned}
    \end{equation*}

    which implies that $Y = \hat{\Lambda}^{1/2} V^t$

    \section{Miscellaneous proofs}

    \subsection{Proof of the relationship between the norm and the trace} \label{norm_trace}

    \begin{equation*}
        \begin{aligned}
            \lVert A \rVert^2 = \sum_{i=1}^n \sum_{j=1}^n A_{ij}^2
            \hspace{1cm}
            Tr(A) = \sum_{i=1}^n A_{ii}
        \end{aligned}
    \end{equation*}\\

    Concider a matrix B that is computed as $A^tA$

    \begin{equation*}
        B = A^tA
    \end{equation*}

    Then an element of B can be computed as

    \begin{equation*}
        B_{ij} = \sum_{k=1}^n A_{ik}^t A_{kj} = \sum_{k=1}^n A_{ki} A_{kj} = (A^tA)_{ij}
    \end{equation*}

    Now, the trace of B will be

    \begin{equation*}
        \begin{aligned}
            Tr(B) &= T(A^tA)\\
                  &= \sum_{i=1}^n (A^tA)_{ii}\\
                  &= \sum_{i=1}^n \sum_{j=1}^n A_{ij}^t A_{ji}\\
                  &= \sum_{i=1}^n \sum_{j=1}^n A_{ji} A_{ji}\\
                  &= \sum_{i=1}^n \sum_{j=1}^n A_{ji}^2 \\
                  &= \sum_{i=1}^n \sum_{j=1}^n A_{ij}^2 \\
                  &= \lVert A \rVert^2
        \end{aligned}
    \end{equation*}

    Hence

    \begin{equation*}
        \lVert A \rVert^2 = Tr(A^tA)
    \end{equation*}

    \subsection{Proof of the circular transformation of the trace} \label{circular_trace}

    Consider a matrix M which is the multiplication of matrices B and C

    \begin{equation*}
        M = BC
    \end{equation*}

    Then an element of M is computed as

    \begin{equation*}
        M_{ij} = \sum_{k=1}^n B_{ik} C_{kj} = BC_{ij}
    \end{equation*}

    Now, the trace of M will be

    \begin{equation*}
        \begin{aligned}
            Tr(M) &= Tr(BC)\\
                  &= \sum_{i=1}^n BC_{ii}\\
                  &= \sum_{i=1}^n \sum_{j=1}^n B_{ij} C_{ji}\\
                  &= \sum_{i=1}^n \sum_{j=1}^n C_{ij} B_{ji}\\
                  &= \sum_{i=1}^n CB_{ii}\\
                  &= Tr(CB)
        \end{aligned}
    \end{equation*}

    hence

    \begin{equation*}
        Tr(BC) = Tr(CB)
    \end{equation*}

    \subsection{Relationship between transpose and inverse of an othogonal matrix}
    \label{orthogonal}

    Consider an orthogonal matrix $Q$,

    \[
        Q =
        \begin{bmatrix}
             |  &        &  | \\
            q_1 & \cdots & q_n\\
             |  &        &  | \\
        \end{bmatrix}
    \]

    Since is orthogonal the following expression holds

    \[
        q_i q_j = 
        \begin{cases} 
           0 & i \neq j \\
           \lVert q_i \rVert & i = j 
        \end{cases}
    \]

    And then

    \[
        Q^t Q =
        \begin{bmatrix}
             - q_1^t - \\
               \vdots  \\
             - q_n^t - \\
        \end{bmatrix}
        \begin{bmatrix}
             |  &        &  | \\
            q_1 & \cdots & q_n\\
             |  &        &  | \\
        \end{bmatrix}
    \]
    \[
        =
        \begin{bmatrix}
            q_1^t q_1 & \cdots & q_1^t q_n \\
              \vdots  & \ddots &  \vdots   \\
            q_n^t q_1 & \cdots & q_n^t q_n \\
        \end{bmatrix}
        =
        \begin{bmatrix}
          \lVert q_1 \rVert &         0         & \cdots &         0    \\
                  0         & \lVert q_2 \rVert & \cdots &         0    \\
               \vdots       &       \vdots      & \ddots &       \vdots \\
                  0         &         0         & \cdots & \lVert q_n \rVert \\
        \end{bmatrix}
    \]

    Which implies that if all the vectors of $Q$ have a length of one ($Q$ is
    orthonormal), then

    \begin{equation*}
        Q^t Q = I
    \end{equation*}

    And if $Q$ is also sqare, then

    \begin{equation*}
        Q^t = Q^{-1}
    \end{equation*}

    \subsection{Spectral decomposition for symmetric matrices} \label{spectral_decomp}

    Consider a symmetric matrix $A$, then this matrix will have $n$ orthogonal
    eigenvalues. Then build a new matrix $S$ that will contain the eigenvalues of
    $A$ as columns.\\

    Now from the definition of eigenvalues and eigenvectors

    \begin{equation*}
        A \mu = \lambda \mu
    \end{equation*}

    If $\mu$ is replaced with $S$ the result would be

    \[
        A S = A 
        \begin{bmatrix}
              |   &        &   |  \\
            \mu_1 & \cdots & \mu_n\\
              |   &        &   |  \\
        \end{bmatrix}
        =
        \begin{bmatrix}
             & - x_1^t - &\\
             &  \vdots   &\\
             & - x_n^t - &\\
        \end{bmatrix}
        \begin{bmatrix}
              |   &        &   |  \\
            \mu_1 & \cdots & \mu_n\\
              |   &        &   |  \\
        \end{bmatrix}\\
    \]
    \[
        =
        \begin{bmatrix}
            x_1^t \mu_1 & \cdots & x_1^t \mu_n\\
               \vdots   & \ddots &   \vdots   \\
            x_n^t \mu_1 & \cdots & x_n^t \mu_n\\
        \end{bmatrix}
        =
        \begin{bmatrix}
                    |       &        &         |      \\
            \lambda_1 \mu_1 & \cdots & \lambda_n \mu_n\\
                    |       &        &         |      \\
        \end{bmatrix}\\
    \]
    \[
        =
        \begin{bmatrix}
               |  &        &   |  \\
            \mu_1 & \cdots & \mu_n\\
               |  &        &   |  \\
        \end{bmatrix}\\
        \begin{bmatrix}
            \lambda_1 &      0    & \cdots &   0 \\
                0     & \lambda_2 & \cdots &   0 \\
              \vdots  &   \vdots  & \ddots & \vdots \\
                0     &      0    & \cdots & \lambda_n \\
        \end{bmatrix}\\
    \]

    So, if a new matrix $\Lambda$ is define as a diagonal matrix that contains
    the eigenvalues of $A$, then the computation can be expressed as

    \begin{equation*}
        A S = S \Lambda
    \end{equation*}

    And, since $S$ is invertible

    \begin{equation*}
        A = S \Lambda S^{-1}
    \end{equation*}

    Finally, as shown in section \ref{orthogonal}, the inverse of an orthonormal
    matrix is its traspose, then A can be expressed as

    \begin{equation*}
        A = S \Lambda S^t
    \end{equation*}

\section{Algorithm implementation}

\section{Results}

\section{Conclusions}

\end{document}

### 2018-11-01 DimReduction/Lle/Documentation/Article_lle.tex
Adding first part of introduction to Lle documentation

@@
 \usepackage{amssymb}
+%\usepackage{babelbib}
 
@@
 \begin{abstract}
-    We implemented the weighted data representation based in KNN and minimization
-    of the euclidean distance between representation. Finally the local embedding
+    We present a detailed implementation of the local linear embeding
+    algorithm from both a theorical and practical approximation. In this work the
+    dimensionality reduction is achieved by a two step procedure. First a
+    weighted data representation based in $K$ nearest neighbors is calculated, and
+    then a minimization of distance between the data representation and a low
+    dimensionality configuration is conducted. Finally the resulting algorithm
     is tested on 3-dimensional manifolds.
@@
 
+Exploratory data analysis and visualization are key issues in a lot of areas
+in science. Gather knowledge about the structure and caracteristics of data
+can lead to great improves in the process of solving problems in science. But a
+good analysis of multidimensinal data can be really hard to achive, and the
+consecuences of mistaken assumptions about the data properties can produce
+dysfunctional solutions.
+
+\vspace{0.25cm}
+
+An approximation to solve this problems are dimensionality reduction techniques.
+These methods provide a way to discover compact representations of
+multidimensional datasets in a low dimensional space. This approaches allow the
+analysis process to be carried out in an easier way and help to reduce the
+probability of getting erroneous results.
+
+\vspace{0.25cm}
+
+However, there are still some issues related to the usage of these methods.
+Assumptions like the use of euclidean distances as a measure in the
+multidimensional space implies that the data is homogeneous or isotropic, and
+this is really often not the case in many real world datasets
+\cite{homogeneous1}, \cite{homogeneous2}. Because of this, the usage of other
+distance metrics is a good option to increase the performance of these approches
+\cite{dist1}, \cite{dist2}.
+
+\vspace{0.25cm}
+
+An appropriate processing method for dimentionality reduction in the case of
+a nonlinear approach can be manifold learning techniques \cite{manifold1},
+\cite{manifold2}. This algorithms seek to generate nonlinear dimensionality
+reductions by exploiting
+
+% missing part
+\cite{lle}
+
+In this work the linear local embeding algorithm is used to perform a
+dimensionality reduction of a multidimentional dataset by a manifold learning
+technique. The first part of the work will carafully describe each part of the
+mathemathical backgound of the method, then the implementation details of the
+algorithm are shown to prepare a results section, and finally some conclusions
+are exposed.
+
 \section{Theorical framework}
 
-\subsection{First optimization problem}
+    As mention before, the problem that we are trying to solve is to find
+    a low dimensional representation of points in a manifold that is in a high
+    dimensional space. In order to achieve this, two steps are considered. The
+    first one is to find a way to represent each point in the original dataset as
+    a linear combination of its $K$ nearest neighbors. This means that the
+    objective is to find a set of weights that minimize the following
+    expression
 
@@
 
-    \subsection{Case for an arbitrary $i$}
+    Where $N(x_i)_j$ represents the $j$-th neighbor of the point $x_i$. The
+    restriction is imposed to get a non trivial solution and ensure that
+    the construction of the points is convex.\\
 
-    define
+    This problem can be solved by optimizing an arbitrary case $i$. The first
+    step is to define the neighborhood matrix of $x_i$, a vector that will
+    conatin the weights of the combination and a special matrix $e$ that will
+    facilitate the process later.
 
@@
 
-    then
+    Whith these tools the linear combinations of neighbors can be expressed as
 
@@
     \end{equation*}
+
+    Hence the problem can be rewritten as follows
 
@@
 
-    define
+    For the next step the objective is to represent $x_i$ as a vector operation.
+    First a matrix in which each column is the $x_i$ point is constructed with
+    the help of the $e$ vector
 
@@
 
-    Now $x_i$ can be defined as
+    Then using the fact that $W_i$ is used to construct a convex representation
+    of the desired point, $x_i$ can be computed as
 
@@
 
-    Hence
+    Now the problem can be redefined in terms of matrix and vectors operations
 
@@
 
-    Define a matrix $G$ as
+    The next step is to define a matrix to simplify the objective function
 
@@
 
-    Redefine the problem as
+    With this matrix the problem can be rewritten as follows
 
@@
 
-    To solve this problem the Lagrange multipliers are used. First redefine the
-    restriction as
+    An option to solve this problem are Lagrange multipliers. In order to use
+    this method the restriction is redefined
 
@@
 
-    Now
+    Then a new objective function can be constructed
+
+    \begin{equation*}
+            \mathcal{L}(W_i, \lambda) = W_i^t G W_i - \lambda (e^t W_i - 1)
+    \end{equation*}
+
+    In order to optimize this function its derivative is calculated and then is
+    set to 0
 
@@
         \begin{aligned}
-            \mathcal{L}(W_i, \lambda) &= W_i^t G W_i - \lambda (e^t W_i - 1)\\
-            \frac{\delta \mathcal{L}}{\delta W_i} &= 2 G W_i - \lambda e
-        \end{aligned}
-    \end{equation*}
-
-    \begin{equation*}
-        \begin{aligned}
+            \frac{\delta \mathcal{L}}{\delta W_i} &= 2 G W_i - \lambda e\\
             2 G W_i - \lambda e &= 0 \\
@@
 
-    Now if $\lambda$ is wrongly chosen, then the result would be a scaled version
-    of the true value. Then, $\lambda$ can be chosen arbitrarily, and after $W_i$
-    is computed a rescale procedure can be conducted in order to satisfy the
-    restriction.\\
+    At this point, if $\lambda$ is wrongly chosen, then the result would be a
+    scaled version of the true value. Then, $\lambda$ can be chosen
+    arbitrarily. Then, after $W_i$ is computed a rescale procedure can be
+    conducted in order to satisfy the restriction.\\
 
-    For simplicity choose $\lambda = 2$, then
+    A good choice for the variable is $\lambda = 2$. This leads to the
+    cancellation of the fraction and reduces the problem to find the solution of
+    a linear system of equations
 
@@
         \begin{aligned}
-            W_i &= \frac{1}{2} \lambda G^{-1} e = G^{-1} e \\
+            W_i = \frac{1}{2} \lambda G^{-1} & e = G^{-1} e \\
             G W_i &= e \\
@@
 
-    So the problem is reduced to solve a linear system of equations. Now if
-    after build the matrix G, it turns out that is a singular matrix the
-    following computation can be used to make it invertible
+    This implies that matrix $G$ must be invertible, and there is no warranty
+    of that. So, if that is the case, a solution is to recalculate $G$ as
 
     \begin{equation*}
-            G = G + \sigma I
+        G \leftarrow G + \sigma I
     \end{equation*}
 
-    where sigma is a small value.
+    where $\sigma$ is a small value.\\
 
-    \subsection{Second optimization problem}
+    This procedure is used to find all the weight values. Then this weights are
+    used to define a configuration of points in a low dimensional space that
+    is as similar as possible to the original dataset. The objective in this
+    point of the problem is minimize the following expression
 
@@
         \begin{aligned}
-            \underset{y}{\text{min}} \quad \sum_{i=1}^n & \lVert y_i - \sum_{j=1}^n w_{ij} y_j \rVert^2 \\
+            \underset{y}{\text{min}} \quad \sum_{i=1}^n & \lVert y_i - \sum_{j=1}^n w_{ij} N(y_i)_j \rVert^2 \\
             & YY^t = I
@@
 
-    Now, define
+    To solve this problem define
 
@@
             |   &  |  &       & |   \\
-            w_1 & w_2 & \dots & w_n \\
+            W_1 & W_2 & \dots & W_n \\
             |   &  |  &       & |
@@
             0 \\
-            \vdots \\ 
             w_1 \\
@@
             w_k \\
-            \vdots \\ 
             0 \\
@@
     to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
-    else. This means that
+    else.\\
+
+    Using this matrix the linear combinations of points can be calculated as
 
     \begin{equation*}
-        X W_i = \sum_{j=1}^k w_{ij} x_j
+        \sum_{j=1}^k w_{ij} N(y_i)_j = Y W_i
     \end{equation*}
 
-    Rewrite the problem as
+    Which means that the problem can be rewrited as
 
     \begin{equation*}
-            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - \sum_{j=1}^n w_{ij} y_j \rVert^2
+            \underset{y}{\text{min}} \quad \sum_{i=1}^n \lVert y_i - \sum_{j=1}^n w_{ij} N(y_i)_j \rVert^2
             =
@@
 
-    Also define
+    Now,
 
@@
 
+\bibliographystyle{unsrt}
+\bibliography{Article_lle}
+
 \end{document}

### 2018-11-04 DimReduction/Mds/Mds.m
Finishing Mds code

@@ -1,23 +1,38 @@
-% colordef none;
-% 
-% n = 1000;
-% [x y z] = build_swiss_roll(n);
-% data = [x y z];
-% 
-% colors = get_cuadrant_colors(x,z);
-% plot_swiss_roll(x,y,z, colors, [-15 15]);
+rng(10); 
+colordef white;
+ 
+n = 1000;
+[x y z] = build_swiss_roll(n);
+data = [x y z];
 
-n = 10;
+colors = get_cuadrant_colors(x,z);
+plot_swiss_roll(x,y,z, colors, [-15 15]);
+
+k = 14;
 p = 2;
+
+dist_matrix = get_distance(data);
+
+graph_matrix = zeros(n);
+
+for i=1:n
+    x_i = data(i,:);
+    neighbors = k_neighbors(dist_matrix, i, k);
+    graph_matrix(i, neighbors) = dist_matrix(i,neighbors);
+    graph_matrix(neighbors,i) = dist_matrix(i,neighbors);
+end
+
+manifold = graph(graph_matrix);
+D = distances(manifold);
 
 H = eye(n) - ones(n)/n;
 A = -0.5 * D.^2;
 B = H * A * H;
 
-[V, L] = eigs(B,p+3,'LR')
+[V, L] = eigs(B,p+3,'LR');
 L_half = sqrt(L);
 Y_full = V * L_half;
-Y = Y(:,1:p)
+Y = Y_full(:,1:p);
+size(colors)
 
-plot(Y(:,1), Y(:,2), '.')
-text(Y(:,1), Y(:,2), names)
+plot_swiss_roll(Y(:,1),Y(:,1),zeros(n,1), colors, [-45 60])

### 2018-11-04 DimReduction/Lle/Documentation/Article_lle.tex
Finishing theorical framework in Lle documentation

@@
 \usepackage{amssymb}
-%\usepackage{babelbib}
 
@@
 \begin{abstract}
-    We present a detailed implementation of the local linear embeding
-    algorithm from both a theorical and practical approximation. In this work the
+    We present a detailed implementation of the local linear embeding algorithm
+    from both a theoretical and practical approximation. In this work the
     dimensionality reduction is achieved by a two step procedure. First a
-    weighted data representation based in $K$ nearest neighbors is calculated, and
-    then a minimization of distance between the data representation and a low
-    dimensionality configuration is conducted. Finally the resulting algorithm
-    is tested on 3-dimensional manifolds.
+    weighted data representation based in $K$ nearest neighbors is calculated,
+    and then a minimization of distance between the data representation and a
+    low dimensionality configuration is conducted. Finally the resulting
+    algorithm is tested on 3-dimensional manifolds.
 \end{abstract}
@@
 
-An approximation to solve this problems are dimensionality reduction techniques.
+An approximation to solve these problems are dimensionality reduction techniques.
 These methods provide a way to discover compact representations of
@@
 Assumptions like the use of euclidean distances as a measure in the
-multidimensional space implies that the data is homogeneous or isotropic, and
+multidimensional space, implies that the data is homogeneous or isotropic, and
 this is really often not the case in many real world datasets
@@
 \cite{manifold2}. This algorithms seek to generate nonlinear dimensionality
-reductions by exploiting
+reductions by assuming that the data points are samples from a low dimensional
+structure which is embedded in a multidimensional space and hence a simpler
+representation can be constructed. \cite{manifold}
 
-% missing part
-\cite{lle}
+\vspace{0.25cm}
 
-In this work the linear local embeding algorithm is used to perform a
-dimensionality reduction of a multidimentional dataset by a manifold learning
-technique. The first part of the work will carafully describe each part of the
-mathemathical backgound of the method, then the implementation details of the
-algorithm are shown to prepare a results section, and finally some conclusions
-are exposed.
+One example of a manifold learning algorithm is Local Linear Embeding (LLE)
+\cite{lle}. This technique uses the fact that from a local perspective a
+non-linear structure can be aproximated with an euclidean measure. Hence an
+aproximation of a multidimensional manifold in a low dimensional space can be
+conducted by representing small pieces, where the euclidean assumption can hold,
+and then using these pieces together to get a reconstruction as similar as 
+possible to the original manifold.
+
+\vspace{0.25cm}
+
+The objective of this work is to build a theoretical and practical develop of
+the ideas used in the LLE algorithm to build a manifold learning method. For
+this purpose the first part of the work will carafully describe each part of
+the mathemathical backgound of the method, then the implementation details of
+the algorithm are shown to prepare a results section, and finally some
+conclusions are exposed.
 
@@
     a low dimensional representation of points in a manifold that is in a high
-    dimensional space. In order to achieve this, two steps are considered. The
+    dimensional space. A theorical approximation of 
+    the method is developed taking as staring point the description given in
+    \cite{proof}. In this method two steps are considered. The
     first one is to find a way to represent each point in the original dataset as
@@
     This problem can be solved by optimizing an arbitrary case $i$. The first
-    step is to define the neighborhood matrix of $x_i$, a vector that will
-    conatin the weights of the combination and a special matrix $e$ that will
-    facilitate the process later.
+    step is to define the neighborhood matrix of $x_i$, also a vector that will
+    conatin the weights of the combination, and finally a special matrix $e$ that
+    is filled with ones.
 
@@
     At this point, if $\lambda$ is wrongly chosen, then the result would be a
-    scaled version of the true value. Then, $\lambda$ can be chosen
-    arbitrarily. Then, after $W_i$ is computed a rescale procedure can be
+    scaled version of the optimal value. Then, $\lambda$ can be chosen
+    arbitrarily and after $W_i$ is computed, a rescale procedure can be
     conducted in order to satisfy the restriction.\\
 
-    A good choice for the variable is $\lambda = 2$. This leads to the
+    Hence, a good choice for the variable is $\lambda = 2$. This leads to the
     cancellation of the fraction and reduces the problem to find the solution of
@@
 
-    This procedure is used to find all the weight values. Then this weights are
-    used to define a configuration of points in a low dimensional space that
+    This procedure is used to find all of the weights values. Then these weights
+    are used to define a configuration of points in a low dimensional space that
     is as similar as possible to the original dataset. The objective in this
-    point of the problem is minimize the following expression
+    point is to minimize the following expression
 
@@
 
-    To solve this problem define
+    To solve this problem, define
 
     \[
+        \hspace{-0.5cm}
         W =
@@
         where
-        \hspace{0.5cm}
+        \hspace{0.25cm}
         W_i =
@@
 
-    Where $W_i$ is a vector that contains the value of the weight that correspond
+    Here $W_i$ is a vector that contains the value of the weight that correspond
     to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
@@
         \sum_{j=1}^k w_{ij} N(y_i)_j = Y W_i
-    \end{equation*}
+    \end{equation*}\\
 
-    Which means that the problem can be rewrited as
+    Which means that the problem can be rewritten as
 
@@
 
-    Now,
+    \vspace{0.25cm}
+
+    Now, just as before, the objective is to express $y_i$ as vectors operations.
+    To achieve this, define
 
@@
     Which is a vector that has a $1$ in the $i$ position and $0$ anywhere else.
-    With this new vector we can redefine the problem as
+    With this new vector the problem can redefined as
 
@@
     \end{equation*}
+
+    \vspace{0.25cm}
 
@@
 
-    So, this optimization problem can be solved as follows
+    This optimization problem can be solved as follows
 
@@
                                                   &= Y M + Y M - 2 \lambda Y\\
-                                                  &= 2 Y M - 2 \lambda Y \\
+                                                  &= 2 Y M - 2 \lambda Y
+        \end{aligned}
+    \end{equation*}
+
+    Finally, by setting this derivative to 0 the solution of the problem can
+    be founded
+
+    \begin{equation*}
+        \begin{aligned}
             2 Y M - 2 \lambda Y &= 0 \\
@@
 
+    Which is a eigenvector problem and can be solved with classical methods.
+
 \section{Algorithm implementation}
+
+\cite{matlab}
 

### 2018-11-05 DimReduction/Lle/Documentation/Article_lle.tex
Adding implementation section to Lle documentation

@@
 \usepackage{amssymb}
+\usepackage[]{algorithm2e}
 
@@
 
-    Whith these tools the linear combinations of neighbors can be expressed as
+    With these tools the linear combinations of neighbors can be expressed as
 
@@
 
-\cite{matlab}
+For the implementation of the method the programing language MATLAB \cite{matlab}
+was used. This tool is specifically desing for matrix operations, which makes it
+an appropriate option to address the problem in an easy way.
+
+\vspace{0.25cm}
+
+The Algorithm \ref{alg1} shows the seudo-code to compute a low
+dimentional representation of a multidimensional manifold that is represented as
+a matrix of points called data of dimensions $n$x$d$.
+
+\begin{algorithm}
+    $n$ = number of points\\
+    $p$ = output dimensions\\
+    $k$ = number of neighbors\\
+    \vspace{0.5cm}
+    $W$ = zeros($n$)\\
+    \For{i in 1:n}{
+        $x_i$ = data($i$)\\
+        $k_i$ = get\_nearest\_neighbors($x_i$, $k$)\\
+        $V_i$ = data($k_i$)\\
+        $G$ = $(x_i e^t - V_i)^t (x_i e^t - V_i)$\\
+        solve($G w_i = e$)\\
+        $W(i)$ = $w_i$ / sum($w_i$)
+    }
+    \vspace{0.5cm}
+    $M$ = $(I-W) * (I-W)^t$\\
+    $V$ = get\_eigenvectors($M$)\\
+    $Y$ = select\_smaller($p$, $V$)
+
+    \vspace{0.5cm}
+
+    \caption{Computation of low dimensional representation}
+    \label{alg1}
+\end{algorithm}
 

### 2018-11-05 DimReduction/Lle/Lle.m

@@ -1,5 +1,5 @@
 rng(10); 
-colordef none;
+colordef white;
 
 n = 1000;
 [x, y, z] = build_swiss_roll(n);
@@ -14,19 +14,17 @@ p = 2;
 tol=0.001;
 
 W = zeros(n);
+e = ones(k,1);
 for i=1:n
     x_i = data(i,:);
     neighbors = k_neighbors(data, x_i, k);
-
     v_i = data(neighbors,:)';
-    e = ones(k,1);
     x_e = repmat(x_i, k, 1)';
     g = (x_e - v_i)' * (x_e - v_i);
     g = g + eye(k)*tol*trace(g); % regularization
     w_i = g\e;
     w_i = w_i / sum(w_i);
     %norm( v_i*w_i - x_i')
-    
     W(neighbors,i) = w_i;
 end
 
@@ -38,5 +36,4 @@ M = (I-W) * (I-W)';
 [eigvec, ~] = eigs(M,p+1,'SM');
 Y = eigvec(:,2:p+1);
 
-plot_swiss_roll(Y(:,1),Y(:,1),zeros(n,1), colors, [-15 15])
 plot_swiss_roll(Y(:,1),Y(:,1),zeros(n,1), colors, [-0.1 0.1])

### 2018-11-05 DimReduction/Lle/Documentation/Article_lle.tex
Adding results section to Lle documentation

@@
 \usepackage[]{algorithm2e}
+\usepackage{float}
 
@@
 \begin{abstract}
-    We present a detailed implementation of the local linear embeding algorithm
+    This work presents a detailed implementation of the local linear embeding algorithm
     from both a theoretical and practical approximation. In this work the
@@
 
+The algorithm described in the previous secction was tested with 2 manifolds in
+a 3 dimensional space. Figure \ref{swiss} shows a swiss roll dataset that
+represents the manifold that the method is going to reduce, the resulting
+2 dimensional dataset is shown in Figure \ref{swiss_res}.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{images/swiss_roll.png}
+    \caption{Swiss roll dataset}
+    \label{swiss}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{images/swiss_roll_result.png}
+    \caption{Swiss roll dataset reduced}
+    \label{swiss_res}
+\end{figure}
+
+Finally in Figure \ref{sphere} a second manifold in form of a 3 dimensional
+sphere is shown, the result of the method is shown in Figure \ref{sphere_res}
+as a 2 dimensional configuration of points.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{images/sphere.png}
+    \caption{Sphere dataset}
+    \label{sphere}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\linewidth]{images/sphere_result.png}
+    \caption{Sphere dataset reduced}
+    \label{sphere_res}
+\end{figure}
+
 \section{Conclusions}
+
+In this work
 

### 2018-11-05 DimReduction/Mds/Documentation/Article_mds.tex
Adding implementation section to Mds documentation

@@
 \begin{abstract}
-    This work presents an theorical approximation to the problem and provide an
-    algorithmic solution easy and understandable. The approach to the problem is
-    reconsider by the construction of the matrix $D^x$ with KNN. Creating the
-    possibility to improve the results using other ways to calculate the distance
-    matrix
+    This work presents an theorical and practical approximation to the problem
+    of non-linear dimensionality reduction using the multidimensional scaling
+    method and provide an algorithmic implementation easy and understandable. In
+    this approach a distance matrix $D^x$ is constructed to approximate the
+    distance over a manifold by solving shortest path problems over a graph that
+    represents a multidimensional structure. This matrix is used as input to the
+    method, which is tested with 3 dimensional manifolds. This technique gives a
+    way to improve the results using other ways to calculate the distance matrix.
 \end{abstract}
@@
 
+Dimensionality reduction is a technique
+
+\cite{manifold}
+
+\cite{dimension}
+
+\cite{mds}
+
 \section{Theorical framework}
 
-    \subsection{Points from distances}
+    As mentioned before, one of the goals of this work is to create a theorical
+    development of the multidimentional scaling method. This section shows a
+    detail explanation of the therical aspects of the method taking \cite{proof}
+    as a base point.
 
-    The set up of the problem is a matrix of euclidean distances $D$, where each
+    \vspace{0.5cm}
+
+    The set up of the problem is a distance matrix $D^x$, where each
     entry is the square of the distance between two points in a $d$ dimensional
-    space.
+    space. The goal of this part is to find a distance matrix in a low
+    dimensional space that is as similar as possible to the multidimensional one.
+    This means that the problem can be expressed in mathematical terms as the
+    following optimization problem
+
+    \begin{equation*}
+        \begin{aligned}
+            \underset{y}{\text{min}}  & \sum_{i=1}^n \sum_{j=1}^n ( d^{(x)}_{ij} - d^{(x)}_{ij} )^2 \\
+            & d_{ij}^{(x)} = \lVert x_i - x_j \rVert\\
+            & d_{ij}^{(y)} = \lVert y_i - y_j \rVert
+        \end{aligned}
+    \end{equation*}\\
+
+    Since each entry of the matrix $D^x$ can be expressed as
+    $d_{ij}^{(x)} = x_i^t x_j$, the problem can be rewritten
+
+    \begin{equation*}
+        \underset{y}{\text{min}} \sum_{i=1}^n \sum_{j=1}^n ( x_i^t x_j - y_i^t y_j )^2 \\
+    \end{equation*}
+
+    For simplicity it can also be expressed in terms of matrix opperations
+
+    \begin{equation*}
+        \underset{Y}{\text{min}} \quad \lVert X^t X - Y^t Y \rVert^2 \\
+    \end{equation*}
+
+    In section \ref{norm_trace}, it was shown that the norm of a matrix can be
+    transformed to a trace, so the problem can be redefined once again
+
+    \begin{equation*}
+        \begin{aligned}
+            A = X^tX &- Y^tY\\
+            \underset{Y}{\text{min}} \quad \lVert X^t X - Y^t Y \rVert^2
+            &=
+            \underset{Y}{\text{min}} \quad \lVert A \rVert^2\\
+            &=
+            \underset{Y}{\text{min}} \quad Tr(A^tA) \\
+        \end{aligned}
+    \end{equation*}
+
+    Now, if the transpose of the matrix $A$ is considered, the result would be
+
+    \begin{equation*}
+        \begin{aligned}
+            A^t &= (X^tX - Y^tY)^t\\
+                &= (X^tX)^t - (Y^tY)^t\\
+                &= X^tX - Y^tY\\
+                &= A
+        \end{aligned}
+    \end{equation*}
+
+    which means that $A^tA = AA = A^2$, and hence
+
+    \begin{equation*}
+        \begin{aligned}
+            \underset{Y}{\text{min}} \quad Tr(A^tA)
+            &=
+            \underset{Y}{\text{min}} \quad Tr(A^2)\\
+            &=
+            \underset{Y}{\text{min}} \quad Tr[(X^tX - Y^tY)^2]
+        \end{aligned}
+    \end{equation*}
+
+    In section \ref{spectral_decomp}, it was shown that $X^tX$ and $Y^tY$ can be
+    expresed as
+
+    \begin{equation*}
+        X^tX = V\Lambda V^t \hspace{1cm} Y^tY = Q \hat{\Lambda} Q^t
+    \end{equation*}
+
+    Then, the problem can be transformed
+
+    \begin{equation*}
+        \underset{Y}{\text{min}} \quad Tr[(X^tX - Y^tY)^2]
+        =
+        \underset{Q,\hat{\Lambda}}{\text{min}} \quad Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]
+    \end{equation*}
+
+    Now, since $V^tV = I = VV^t$, then using the proof presented in section
+    \ref{circular_trace} the following propertie can be deducted
+
+    \begin{equation*}
+        \begin{aligned}
+            A =& V\Lambda V^t - Q \hat{\Lambda} Q^t\\
+            Tr[(V^tAV)^2] &= Tr[V^tAVV^tAV]\\
+                          &= Tr[V^tAIAV]\\
+                          &= Tr[AIAVV^t]\\
+                          &= Tr[AIAI]\\
+                          &= Tr[AA]\\
+                          &= Tr[A^2]\\
+        \end{aligned}
+    \end{equation*}
+
+    Which implies that
+
+    \begin{equation*}
+            \hspace{-1cm}
+        \begin{aligned}
+            Tr[(V^tAV)^2] &= Tr[A^2]\\
+            Tr[(V^t(V\Lambda V^t - Q \hat{\Lambda} Q^t)V)^2] &= Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]\\
+            Tr[(V^tV\Lambda V^tV - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
+            Tr[(I\Lambda I - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
+            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
+        \end{aligned}
+    \end{equation*}
+
+    At this point, a new matrix $G$ can be defined to simplify the expression
+
+    \begin{equation*}
+        \begin{aligned}
+            G &= V^tQ\\
+            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &= Tr[(\Lambda - G \hat{\Lambda} G^t)^2]\\
+        \end{aligned}
+    \end{equation*}
+
+    Hence the problem can be redefined one more time
+
+    \begin{equation*}
+        \hspace{-0.5cm}
+        \begin{aligned}
+        \underset{Q,\hat{\Lambda}}{\text{min}} \quad Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]
+        &=
+        \underset{G,\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
+        \end{aligned}
+    \end{equation*}
+
+    Now, using that knowledge, the main expression can be expanded
+
+    \begin{equation*}
+        \begin{aligned}
+        Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
+        &=
+        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + (G \hat{\Lambda} G^t)^2]\\
+        =
+        &Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} G^t G \hat{\Lambda} G^t]\\
+        \end{aligned}
+    \end{equation*}
+
+    But, knowing that $Q^tQ = I = QQ^t$, the following properties can be developed
+
+    \begin{equation*}
+        \begin{aligned}
+            G^tG &= (V^tQ)^t V^tQ\\
+                 &= (Q^tV) V^tQ\\
+                 &= Q^tQ\\
+                 &= I
+        \end{aligned}
+    \end{equation*}
+
+    \begin{equation*}
+        \begin{aligned}
+            GG^t &= V^tQ (V^tQ)^t\\
+                 &= V^tQ (Q^tV)\\
+                 &= V^tV\\
+                 &= I\\ \\
+        \end{aligned}
+    \end{equation*}
+
+    So, the following transformations can be made to the problem
+
+    \begin{equation*}
+        \begin{aligned}
+        Tr[(\Lambda - G \hat{\Lambda} G^t G)^2 ]
+        &=
+        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} \hat{\Lambda} G^t]\\
+        =
+        Tr[&\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda}^2 G^t]\\
+        =
+        Tr[&\Lambda^2]- 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[G \hat{\Lambda}^2 G^t]\\
+        =
+        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2 G^tG]\\
+        =
+        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2]\\
+        =
+        Tr[&\Lambda^2] + Tr[ \hat{\Lambda}^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t]\\
+        \end{aligned}
+    \end{equation*}
+
+    Which takes us to the \textbf{DARK PART!}\\
+
+    We know that $\frac{\delta Tr[\Lambda G \hat{\Lambda} G^t]}{\delta G} = 2\Lambda G \hat{\Lambda}$\\\\
+    We also know that $\frac{\delta Tr(\Lambda^2)}{\delta G} = 0 = \frac{\delta Tr(\hat{\Lambda}^2)}{\delta G}$\\\\
+    Finally we know that the answer for a fixed $\hat{\Lambda}$ is $G=I$\\
+
+    \textbf{END OF DARK PART}\\
+
+    This result implies that
+
+    \begin{equation*}
+        \begin{aligned}
+        G = I = V^tQ\\
+        V^tQ = I\\
+        V(V^tQ) = VI\\
+        Q = V
+        \end{aligned}
+    \end{equation*}
+
+    Now, coming back to the original problem this information can be used to
+    simplify the objective function
+
+    \begin{equation*}
+        \begin{aligned}
+            \underset{G,\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
+            &=
+            \underset{\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - \hat{\Lambda})^2]
+        \end{aligned}
+    \end{equation*}
+
+    To minimize the resulting expression, the first $p$ eigenvalues of $\Lambda$
+    can be used to define $\hat{\Lambda}$. And finally the solution to the
+    problem can be found by the following procedure
+
+    \begin{equation*}
+        \begin{aligned}
+            Y^tY &= Q \hat{\Lambda} Q^t\\
+            &=
+            V \hat{\Lambda} V^t\\
+            &=
+            V \hat{\Lambda}^{1/2} \hat{\Lambda}^{1/2} V^t\\
+        \end{aligned}
+    \end{equation*}
+
+    which implies that $Y = \hat{\Lambda}^{1/2} V^t$.
+
+    \vspace{0.25cm}
+
+    This method gives a way to find a matrix $D$ that encodes the distances
+    between points that are in a low dimensional space. Now a procedure that
+    takes a distance matrix as input and finds a corresponding configuration
+    of points is described. The distance matrix is defined as follows
 
@@
 
-    The unknown of the problem is a matrix $X$ that cotains the position of the
+    The unknown of this problem is a matrix $X$ that cotains the position of the
     $n$ points in the $d$ dimensional space.
@@
 
-    But there is a restriction to make sure that the matrix is centered at 0.
+    But there is a restriction to make sure that the matrix is centered at 0 and
+    avoid indeterminations due to the fact that any point can be choosed as center.
 
@@
 
-    \subsection{Solution}
-
-    First rewrite each entry of matrix $D$ in a vector form
+    the first step to solve the problem is rewrite each entry of matrix $D$ in
+    a vector form
 
@@
 
-    Then expand distance $d_{ij}^2$
+    Then, each component $d_{ij}^2$ can be expanded as follows
 
@@
     The goal is to express all the dot products of $x$ vectors in terms of the
-    $d_{ij}^2$ distances. Consider the following expression
+    $d_{ij}^2$ distances. To achieve this goal, consider the following expression
 
@@
 
-    Remember that
+    But, using the restriction, one term can be canceled
 
@@
 
-    Finally, the first approximation to the problem can be constructed
+    Finally, the first approximation to the solution of the problem can be
+    constructed
 
@@
 
-    Notice that
-
-    \begin{equation*}
-        \sum_{i=1}^n x_i^t x_i = \sum_{j=1}^n x_j^t x_j
-    \end{equation*}
-
-    So
+    But this expression has a term that can be treated as duplicated, so the
+    following propertie can be applied to simplify the expression
 
@@
         \begin{aligned}
-            x_i^t x_j&=
-            \frac{1}{2} \left[ \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - d_{ij}^2 + \frac{1}{n} \sum_{i=1}^n d_{ij}^2 - \frac{2}{n} \sum_{i=1}^n x_i^t x_i \right]\\
+            \sum_{i=1}^n x_i^t x_i &= \sum_{j=1}^n x_j^t x_j\\
+            x_i^t x_j =
+            \frac{1}{2} \Bigg [ \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - d_{ij}^2 + &\frac{1}{n} \sum_{i=1}^n d_{ij}^2 - \frac{2}{n} \sum_{i=1}^n x_i^t x_i \Bigg ]\\
         \end{aligned}
@@
 
-    Now the expression can be completed
+    Now the goal is completed, and the dot product between the $X$ components
+    can be expressed in terms of distance operations
+
     \begin{equation*}
@@
 
-    This computation can be expressed in matrix form to simplify the formula.
-    First define
+    This computation can also be expressed in matrix form to simplify the formula.
+    First the following definitions must take place
 
@@
 
-    Now a matrix $B$ can be defined as
+    Then, a matrix $B$ can be defined as
 
@@
     This formula gives a way to compute all the dot products between the points
-    that makes up the matrix $X$. Because of this the matrix $B$ can also
+    that makes up the matrix $X$. Because of this, the matrix $B$ can also
     be expressed as
@@
 
-    Which gives a way to compute all the points of the matrix $X$.
+    Which gives a way to compute all the points of the matrix $X$ finding a low
+    dimensional representation of the original dataset.
 
-    \section{Mds optimization problem}
+\section{Algorithm implementation}
 
-    \begin{equation*}
-        \begin{aligned}
-            \underset{y}{\text{min}}  & \sum_{i=1}^n \sum_{j=1}^n ( d^{(x)}_{ij} - d^{(x)}_{ij} )^2 \\
-            & d_{ij}^{(x)} = \lVert x_i - x_j \rVert\\
-            & d_{ij}^{(y)} = \lVert y_i - y_j \rVert
-        \end{aligned}
-    \end{equation*}\\
+\begin{algorithm}
+    $n$ = number of points\\
+    $p$ = output dimensions\\
+    $k$ = number of neighbors\\
+    \vspace{0.5cm}
+    $D_e$ = compute\_euclidean\_distance(data)\\
+    $G$ = zeros($n$) //graph adjacency matrix\\
+    \vspace{0.5cm}
+    \For{i in 1:n}{
+        $x_i$ = data($i$)\\
+        $k_i$ = get\_nearest\_neighbors($x_i$, $k$)\\
+        $G(i,k_i)$ = $D_e$($i$, $k_i$)\\
+        $G(k_i,i)$ = $D_e$($i$, $k_i$)\\
+    }
+    \vspace{0.5cm}
+    $H$ = $I - \frac{1}{n} e e^t$\\
+    $D^x$ = compute\_distances($G$)\\
+    $A$ = -$\frac{1}{2}$ pow\_elementwise($D^x$,2)\\
+    $B$ = $HAH$\\
+    $V$ = get\_eigenvectors($B$)\\
+    $\Lambda$ = get\_eigenvalues($B$)\\
+    $V$, $\Lambda$ = select\_smallers($p$, $V$, $\Lambda$)\\
+    $Y$ = $\Lambda^{1/2} V^t$
 
-    if $d_{ij}^{(x)}$ es euclidean, then $d_{ij}^{(x)} = x_i^t x_j$, and hence
+    \vspace{0.5cm}
 
-    \begin{equation*}
-        \underset{y}{\text{min}} \sum_{i=1}^n \sum_{j=1}^n ( x_i^t x_j - y_i^t y_j )^2 \\
-    \end{equation*}
+    \caption{Computation of low dimensional representation}
+    \label{alg1}
+\end{algorithm}
 
-    or in matrix form
+\cite{matlab}
 
-    \begin{equation*}
-        \underset{Y}{\text{min}} \quad \lVert X^t X - Y^t Y \rVert^2 \\
-    \end{equation*}
+\section{Results}
 
-    As shown in section \ref{norm_trace}, the norm of a matrix can be transformed
-    to a trace, so the problem can be redefined as
+\section{Conclusions}
 
-    \begin{equation*}
-        \begin{aligned}
-            A = X^tX &- Y^tY\\
-            \underset{Y}{\text{min}} \quad \lVert X^t X - Y^t Y \rVert^2
-            &=
-            \underset{Y}{\text{min}} \quad \lVert A \rVert^2\\
-            &=
-            \underset{Y}{\text{min}} \quad Tr(A^tA) \\
-        \end{aligned}
-    \end{equation*}\\
-
-    Now
-
-    \begin{equation*}
-        \begin{aligned}
-            A^t &= (X^tX - Y^tY)^t\\
-                &= (X^tX)^t - (Y^tY)^t\\
-                &= X^tX - Y^tY\\
-                &= A
-        \end{aligned}
-    \end{equation*}
-
-    Which means that $A^tA = AA = A^2$, and hence
-
-    \begin{equation*}
-        \begin{aligned}
-            \underset{Y}{\text{min}} \quad Tr(A^tA)
-            &=
-            \underset{Y}{\text{min}} \quad Tr(A^2)\\
-            &=
-            \underset{Y}{\text{min}} \quad Tr[(X^tX - Y^tY)^2]
-        \end{aligned}
-    \end{equation*}\\
-
-    now as shown in section \ref{spectral_decomp}, $X^tX$ and $Y^tY$ can be
-    expresed as
-
-    \begin{equation*}
-        X^tX = V\Lambda V^t \hspace{1cm} Y^tY = Q \hat{\Lambda} Q^t
-    \end{equation*}
-
-    Then
-
-    \begin{equation*}
-        \underset{Y}{\text{min}} \quad Tr[(X^tX - Y^tY)^2]
-        =
-        \underset{Q,\hat{\Lambda}}{\text{min}} \quad Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]
-    \end{equation*}
-
-    We know that $V^tV = I = VV^t$, then using the proof presented in section \ref{circular_trace}
-
-    \begin{equation*}
-        \begin{aligned}
-            A =& V\Lambda V^t - Q \hat{\Lambda} Q^t\\
-            Tr[(V^tAV)^2] &= Tr[V^tAVV^tAV]\\
-                          &= Tr[V^tAIAV]\\
-                          &= Tr[AIAVV^t]\\
-                          &= Tr[AIAI]\\
-                          &= Tr[AA]\\
-                          &= Tr[A^2]\\
-        \end{aligned}
-    \end{equation*}\\
-
-    Which implies that
-
-    \begin{equation*}
-        \begin{aligned}
-            Tr[(V^tAV)^2] &= Tr[A^2]\\
-            Tr[(V^t(V\Lambda V^t - Q \hat{\Lambda} Q^t)V)^2] &= Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]\\
-            Tr[(V^tV\Lambda V^tV - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
-            Tr[(I\Lambda I - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
-            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &=\\
-        \end{aligned}
-    \end{equation*}
-
-    Now
-
-    \begin{equation*}
-        \begin{aligned}
-            G &= V^tQ\\
-            Tr[(\Lambda - V^tQ \hat{\Lambda} Q^tV)^2] &= Tr[(\Lambda - G \hat{\Lambda} G^t)^2]\\
-        \end{aligned}
-    \end{equation*}\\
-
-    Which means that the problem can be redefined one more time
-
-    \begin{equation*}
-        \begin{aligned}
-        \underset{Q,\hat{\Lambda}}{\text{min}} \quad Tr[(V\Lambda V^t - Q \hat{\Lambda} Q^t)^2]
-        &=
-        \underset{G,\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
-        \end{aligned}
-    \end{equation*}\\
-
-    Now
-
-    \begin{equation*}
-        \begin{aligned}
-        Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
-        &=
-        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + (G \hat{\Lambda} G^t)^2]\\
-        &=
-        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} G^t G \hat{\Lambda} G^t]\\
-        \end{aligned}
-    \end{equation*}
-
-    But (Knowing that $Q^tQ = I = QQ^t$)
-
-    \begin{equation*}
-        \begin{aligned}
-            G^tG &= (V^tQ)^t V^tQ\\
-                 &= (Q^tV) V^tQ\\
-                 &= Q^tQ\\
-                 &= I\\ \\
-            GG^t &= V^tQ (V^tQ)^t\\
-                 &= V^tQ (Q^tV)\\
-                 &= V^tV\\
-                 &= I\\ \\
-        \end{aligned}
-    \end{equation*}
-
-    Hence
-
-    \begin{equation*}
-        \begin{aligned}
-        Tr[(\Lambda - G \hat{\Lambda} G^t G)^2 ]
-        &=
-        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} \hat{\Lambda} G^t]\\
-        =
-        Tr[&\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda}^2 G^t]\\
-        =
-        Tr[&\Lambda^2]- 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[G \hat{\Lambda}^2 G^t]\\
-        =
-        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2 G^tG]\\
-        =
-        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2]\\
-        =
-        Tr[&\Lambda^2] + Tr[ \hat{\Lambda}^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t]\\
-        \end{aligned}
-    \end{equation*}
-
-    \subsection{Dark part}
-
-    We know that $\frac{\delta Tr[\Lambda G \hat{\Lambda} G^t]}{\delta G} = 2\Lambda G \hat{\Lambda}$\\\\
-    We also know that $\frac{\delta Tr(\Lambda^2)}{\delta G} = 0 = \frac{\delta Tr(\hat{\Lambda}^2)}{\delta G}$\\\\
-    Finally we know that the answer for a fixed $\hat{\Lambda}$ is $G=I$\\ \\
-
-    Which implies that
-
-    \begin{equation*}
-        \begin{aligned}
-        G = I = V^tQ\\
-        V^tQ = I\\
-        V(V^tQ) = VI\\
-        Q = V
-        \end{aligned}
-    \end{equation*}
-
-    Now, coming back to the original problem
-
-    \begin{equation*}
-        \begin{aligned}
-            \underset{G,\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
-            &=
-            \underset{\hat{\Lambda}}{\text{min}} \quad Tr[(\Lambda - \hat{\Lambda})^2]
-        \end{aligned}
-    \end{equation*}
-
-    So we can take the first $p$ eigenvalues of $\Lambda$ to define $\hat{\Lambda}$,
-    and finally
-
-    \begin{equation*}
-        \begin{aligned}
-            Y^tY &= Q \hat{\Lambda} Q^t\\
-            &=
-            V \hat{\Lambda} V^t\\
-            &=
-            V \hat{\Lambda}^{1/2} \hat{\Lambda}^{1/2} V^t\\
-        \end{aligned}
-    \end{equation*}
-
-    which implies that $Y = \hat{\Lambda}^{1/2} V^t$
-
-    \section{Miscellaneous proofs}
+    \section{Apendix}
 
@@
 
-    Concider a matrix B that is computed as $A^tA$
+    Consider a matrix B that is computed as $A^tA$
 
@@
 
-    Then an element of B can be computed as
+    Then, an element of B can be computed as
 
@@
 
-\section{Algorithm implementation}
-
-\section{Results}
-
-\section{Conclusions}
+\bibliographystyle{unsrt}
+\bibliography{Article_mds}
 

### 2018-11-05 DimReduction/Mds/Documentation/Article_mds.tex
Adding result section to Mds documentation

@@
     As mentioned before, one of the goals of this work is to create a theorical
-    development of the multidimentional scaling method. This section shows a
+    development of the multidimentional scaling algorithm. This section shows a
     detail explanation of the therical aspects of the method taking \cite{proof}
-    as a base point.
+    as base point.
 
@@
     Since each entry of the matrix $D^x$ can be expressed as
-    $d_{ij}^{(x)} = x_i^t x_j$, the problem can be rewritten
+    $d_{ij}^{(x)} = x_i^t x_j$, then the problem can be rewritten as
 
@@
     Now, coming back to the original problem this information can be used to
-    simplify the objective function
+    simplify the objective function as follows
 
@@
 
-    which implies that $Y = \hat{\Lambda}^{1/2} V^t$.
+    which implies that $Y = \hat{\Lambda}^{1/2} V^t$. Which solves the problem.
 
@@
 
-    This method gives a way to find a matrix $D$ that encodes the distances
-    between points that are in a low dimensional space. Now a procedure that
+    This method gives a way to find a matrix $D^x$ that encodes the distances
+    between points that are in a low dimensional space. Now, a procedure that
     takes a distance matrix as input and finds a corresponding configuration
-    of points is described. The distance matrix is defined as follows
+    of points is described.
+
+    \vspace{0.25cm}
+
+    First, the distance matrix is defined as follows
 
@@
     The unknown of this problem is a matrix $X$ that cotains the position of the
-    $n$ points in the $d$ dimensional space.
+    $n$ points in a $d$ dimensional space.
 
@@
 
-    the first step to solve the problem is rewrite each entry of matrix $D$ in
+    To solve the problem the first step is rewrite each entry of matrix $D$ in
     a vector form
@@
     The goal is to express all the dot products of $x$ vectors in terms of the
-    $d_{ij}^2$ distances. To achieve this goal, consider the following expression
+    $d_{ij}^2$ distances. To achieve this goal, consider the following
+    expressions
 
@@
 
-    But, using the restriction, one term can be canceled
+    But, using the restriction, one term must be canceled
 
@@
 
-    Now in the other hand, the sum over $j$ would be
+    Now, in the other hand, the sum over $j$ would be
 
@@
             x_i^t x_j &= \frac{1}{2} (x_i^t x_i - d_{ij}^2 + x_j^t x_j)\\
-            \text{where}\\
+        \end{aligned}
+    \end{equation*}
+
+    Where the dot products of the left side can be replaced by the expressions
+    shown below
+
+    \begin{equation*}
+        \begin{aligned}
             x_i^t x_i &= \frac{1}{n} \sum_{j=1}^n d_{ij}^2 - \frac{1}{n} \sum_{j=1}^n x_j^t x_j\\
@@
 
-    But this expression has a term that can be treated as duplicated, so the
+    But this expression have a term that can be treated as duplicated, so the
     following propertie can be applied to simplify the expression
@@
 
-    This is almost expressed in terms of the distances, the only problem is the
-    last term. To express that part consider the following expression
+    This is almost expressed in terms of distance operations, the only problem
+    is the last term. To express that part consider the following expression
 
@@
 
-    Now the goal is completed, and the dot product between the $X$ components
+    Now the goal is completed, and the dot product between the $x$ vectors
     can be expressed in terms of distance operations
@@
 
-    Which gives a way to compute all the points of the matrix $X$ finding a low
+    Which gives a way to compute all the points of the matrix $X$, finding a low
     dimensional representation of the original dataset.
@@
 \section{Algorithm implementation}
+
+In this section the implementation details of the method are exposed. The method
+described in the previous section was developed using the Matlab programming
+language \cite{matlab}. Algorithm \ref{alg1} shows a seudo-code of the
+implemented method.
 
@@
     }
+    $D^x$ = compute\_distances($G$)\\
     \vspace{0.5cm}
+    $e$ = vector\_ones($n$)\\
     $H$ = $I - \frac{1}{n} e e^t$\\
-    $D^x$ = compute\_distances($G$)\\
     $A$ = -$\frac{1}{2}$ pow\_elementwise($D^x$,2)\\
@@
 
-\cite{matlab}
+This algorithm is divided in two main parts. The first one computes the initial
+distance matrix, that represents the geodesic distance in the multidimensional
+manifold. This computation is achieved by solving the shortest path problem
+between all the pairs of nodes in a graph that represents the manifold. This
+graph is constructed by taking each point as a node, and building an arc from
+each node to its $k$ nearest neighbors. As a last step, the weight of each arc
+is initialized as the euclidean distance between the two nodes connected.
+
+\vspace{0.5cm}
+
+And finally the procedure described in the theorical framework can be executed
+to get a low dimensional representation of the original multidimensional
+manifold.
 
 \section{Results}
+
+The method described in the previuos section was tested with 3 dimensional
+manifolds in order to allow the visualization of the intial manifold and be able
+to compare the quality of the resulting solution. Figure \ref{swiss} shows
+a swiss roll dataset that was used as the multidimentional manifold. Figure
+\ref{swiss_res} shows the result of the devolped method.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{images/swiss_roll.png}
+    \caption{Swiss roll dataset}
+    \label{swiss}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\linewidth]{images/swiss_roll_result.png}
+    \caption{Swiss roll dataset reduced}
+    \label{swiss_res}
+\end{figure}
 
@@
 
-    hence
+    Hence
 

### 2018-11-05 DimReduction/Mds/Documentation/Article_mds.tex
Adding introduction to Mds documentation

@@
 
-Dimensionality reduction is a technique
+Dimensionality reduction is a technique that aims to summarize the information
+of a dataset in the most compact way posible in order to facilitate the analysis
+procedures of the data. These procedures can also be seen as a process to find
+a set of degrees of freedom that econde most of the dataset variability
+\cite{proof}.
 
-\cite{manifold}
+\vspace{0.5cm}
 
+Some of the approximations in these types of methods assume that the data that
+is being analyzed does not exploit all the dimensionality of the space where
+is defined. Due to this fact, another description of the goal that the
+dimensionality reduction techniques have is related to find a meaningful low
+dimensional structure that is hidden in a multidimentional structure.
 \cite{dimension}
 
-\cite{mds}
+\vspace{0.5cm}
+
+A clear example of these type of methods are manifold learning techniques. These
+approximations search for a way to perform a non linear dimensionality reduction,
+which means that they do not assume that the structure of the original dataset
+is encoded based on euclidean distance meassuares. Instead these algorithms try
+to express their relationships in terms of geodesic distances, which are
+distances meassured over the manifold surface. The result of this intend is 
+to preserve the intrinsic structure of the data when the reduction is perfomed,
+allowing to conserve the implicit information that was hidden in the original
+structure of the dataset. \cite{manifold}.
+
+\vspace{0.5cm}
+
+One example of a manifold learning algorithm is called multidimensional
+scaling \cite{mds}, this method addresses the problem by trying to find a low
+dimensional distance matrix that is as close as possible to a multidimensional
+version that encodes the relationship between the points in the original space.
+Then the algorithm uses the founded matrix to build a correspondent configuration
+of points in a low dimentional space that will act as a less complicated
+representation of the original dataset.
+
+\vspace{0.5cm}
+
+This work aims to develop the theorical and practical aspects of the method
+described above, with the objective of build a method that can reduce a
+multidimensional manifold to a low dimensional representation. To achieve this
+in the first part of the work a theorical framework is developed, in order to
+understand all the aspects of the method. Then an implementation section will
+take place, followed by a result section. Later, some conclusions are
+pressented and finally an apendix is added to present some proofs of the
+steps that are used in the theorical framework.
 
@@
     \begin{equation*}
-            \hspace{-1cm}
         \begin{aligned}
@@
             \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n [x_i^t x_i - 2 x_i^t x_j + x_j^t x_j]\\
-            &=
-            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i - \frac{2(0)}{n^2} +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
-            &=
-            \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
-            &=
-            \frac{n}{n^2}\sum_{i=1}^n x_i^t x_i +  \frac{n}{n^2}\sum_{j=1}^n x_j^t x_j\\
-            &=
-            \frac{1}{n}\sum_{i=1}^n x_i^t x_i +  \frac{1}{n}\sum_{j=1}^n x_j^t x_j\\
-            &=
-            \frac{2}{n}\sum_{i=1}^n x_i^t x_i\\
+            =
+            \frac{1}{n^2}&\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i - \frac{2(0)}{n^2} +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
+            =
+            \frac{1}{n^2}&\sum_{i=1}^n \sum_{j=1}^n x_i^t x_i +  \frac{1}{n^2}\sum_{i=1}^n \sum_{j=1}^nx_j^t x_j\\
+            =
+            \frac{n}{n^2}&\sum_{i=1}^n x_i^t x_i +  \frac{n}{n^2}\sum_{j=1}^n x_j^t x_j\\
+            =
+            \frac{1}{n}&\sum_{i=1}^n x_i^t x_i +  \frac{1}{n}\sum_{j=1}^n x_j^t x_j\\
+            =
+            \frac{2}{n}&\sum_{i=1}^n x_i^t x_i\\
         \end{aligned}
@@
     \begin{equation*}
+        \hspace{-0.5cm}
         \begin{aligned}

### 2018-11-06 DimReduction/Lle/Documentation/Article_lle.tex
Adding conclusion section to Lle documentation

@@
 \begin{abstract}
-    This work presents a detailed implementation of the local linear embeding algorithm
-    from both a theoretical and practical approximation. In this work the
-    dimensionality reduction is achieved by a two step procedure. First a
+    This work presents a detailed implementation of the local linear embeding
+    algorithm from both a theoretical and practical approximation. In this work
+    the dimensionality reduction is achieved by a two step procedure. First a
     weighted data representation based in $K$ nearest neighbors is calculated,
@@
 
-In this work
+In this work a theorical and practical develop of the local linear embeding
+algorithm was described. During the process it was found that althought the
+problems started with a quite complicated setup, the development of the problem
+allowed to generate great simplifications that reduced the original problem to
+just solve a linear system of equations and an eigenvector problem. These
+simplifications made possible the creation of a relatively simple algorithm that
+can detect low dimensional manifolds in multidimensional spaces. This statement
+is supported in the results section where 3 dimensional manifolds were
+simplified to dimension 2 but preserving the intrinsic structure of the original
+dataset.
 

### 2018-11-06 DimReduction/Mds/Documentation/Article_mds.tex
Adding conclusions to Mds documentation

@@
 dimensionality reduction techniques have is related to find a meaningful low
-dimensional structure that is hidden in a multidimentional structure.
+dimensional structure that is hidden in a multidimensional structure.
 \cite{dimension}
@@
 Then the algorithm uses the founded matrix to build a correspondent configuration
-of points in a low dimentional space that will act as a less complicated
+of points in a low dimensional space that will act as a less complicated
 representation of the original dataset.
@@
     As mentioned before, one of the goals of this work is to create a theorical
-    development of the multidimentional scaling algorithm. This section shows a
+    development of the multidimensional scaling algorithm. This section shows a
     detail explanation of the therical aspects of the method taking \cite{proof}
@@
 to compare the quality of the resulting solution. Figure \ref{swiss} shows
-a swiss roll dataset that was used as the multidimentional manifold. Figure
+a swiss roll dataset that was used as the multidimensional manifold. Figure
 \ref{swiss_res} shows the result of the devolped method.
@@
 \section{Conclusions}
+
+In this work a theorical and practical development of the multidimensional
+scaling method was permed. During the process it was found that what seemed like
+a fairly complicated problem, can be simplified to simple matrix operations
+with the help of some concepts from linear algebra. Thanks to that theorical
+framework, the construction of a relatively simple algorithm that can detect
+low dimensional manifolds embedded in multidimensional spaces was possible.
+Finally the results shown that this technique has the ability to make
+contributions to the analysis proces of multidimensional datasets by
+simplifying its features but preserving the intrinsic structure.
 
@@
                   &= \sum_{i=1}^n \sum_{j=1}^n A_{ji}^2 \\
-                  &= \sum_{i=1}^n \sum_{j=1}^n A_{ij}^2 \\
-                  &= \lVert A \rVert^2
+                  &= \sum_{i=1}^n \sum_{j=1}^n A_{ij}^2 = \lVert A \rVert^2 \\
         \end{aligned}

### 2018-11-16 DimReduction/Mds/Mds.m
Adding s curve dataset to mds

@@ -1,12 +1,12 @@
 rng(10); 
 colordef white;
  
-n = 1000;
-[x y z] = build_swiss_roll(n);
+n = 1500;
+[x y z] = build_s_curve(n);
 data = [x y z];
 
 colors = get_cuadrant_colors(x,z);
-plot_swiss_roll(x,y,z, colors, [-15 15]);
+plot_swiss_roll(x,y,z, colors, [-2 2]);
 
 k = 14;
 p = 2;
@@ -33,6 +33,5 @@ B = H * A * H;
 L_half = sqrt(L);
 Y_full = V * L_half;
 Y = Y_full(:,1:p);
-size(colors)
 
-plot_swiss_roll(Y(:,1),Y(:,1),zeros(n,1), colors, [-45 60])
+plot_swiss_roll(Y(:,1),Y(:,1),zeros(n,1), colors, [-6 6])

### 2018-11-17 DimReduction/Documentation/poster.tex
Adding global documentation to DimReduction

\documentclass[a0,portrait]{a0poster}

\usepackage{multicol}
\columnsep=100pt
\columnseprule=3pt

\usepackage[svgnames]{xcolor}
\usepackage{graphicx}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{amsfonts, amsmath, amsthm, amssymb}
\usepackage{wrapfig}
\usepackage[]{algorithm2e}
\usepackage{booktabs}

\begin{document}

\begin{minipage}[b]{0.80\linewidth}
\VeryHuge \color{NavyBlue} \textbf{A theorical and practical approximation to manifold learning} \color{Black}\\[2.4cm] % Title
\huge \textbf{Alejandro Salgado G. and O. Luc\'ia Quintero M.}\\[0.5cm] % Author(s)
\LARGE Mathematical Science Department, School of Sciences, Universidad EAFIT\\[0.4cm]
\Large \texttt{asalgad2@eafit.edu.co, oquinte1@eafit.edu.co}\\
\end{minipage}
%
\begin{minipage}[b]{0.2\linewidth}
\includegraphics[width=14cm]{figures/logo.png}\\
\end{minipage}

\begin{multicols}{3}

    \begin{abstract}
    This work presents a theorical and practical approximation to the problem
    of non-linear dimensionality reduction using the local linear embedding
    algorithm (Lle) and the multidimensional scaling method (Mds). In this work
    the Lle algorithm is implemented as a two step procedure. First a weighted
    data representation based in $K$ nearest neighbors is calculated, then a
    minimization of distance between the data representation and a low
    dimensionality configuration is conducted. In the second approach a
    distance matrix $D^x$ is constructed to approximate the distance over a
    manifold by solving shortest path problems over a graph that represents a
    multidimensional structure, then this matrix is used as input to the
    method. Finally the resulting algorithms are tested on 3-dimensional
    manifolds.
    \end{abstract}

    \section*{Introduction}

    Exploratory data analysis and visualization are key issues in a lot of areas
    in science. Gather knowledge about the structure and caracteristics of data
    can lead to great improves in the process of solving problems in science. But a
    good analysis of multidimensinal data can be really hard to achive, and the
    consecuences of mistaken assumptions about the data properties can produce
    dysfunctional solutions.\\

    An approximation to solve these problems are dimensionality reduction techniques.
    However, there are still some issues related to the usage of these methods.
    Assumptions like the use of euclidean distances as a measure in the
    multidimensional space, implies that the data is homogeneous or isotropic, and
    this is really often not the case in many real world datasets
    \cite{homogeneous1}, \cite{homogeneous2}. Because of this, the usage of other
    distance metrics is a good option to increase the performance of these approches
    \cite{dist1}, \cite{dist2}.\\

    An appropriate processing method for dimentionality reduction in the case of
    a nonlinear approach can be manifold learning techniques \cite{manifold1},
    \cite{manifold2}. This algorithms seek to generate nonlinear dimensionality
    reductions by assuming that the data points are samples from a low dimensional
    structure which is embedded in a multidimensional space and hence a simpler
    representation can be constructed. \cite{manifold_lle}\\

    One example of a manifold learning algorithm is Local Linear Embeding (LLE)
    \cite{lle}. This technique uses the fact that from a local perspective a
    non-linear structure can be aproximated with an euclidean measure. Hence, an
    aproximation of a multidimensional manifold in a low dimensional space can be
    conducted by representing small pieces, where the euclidean assumption can hold,
    and then using these pieces together to get a reconstruction as similar as
    possible to the original manifold. Another example is multidimensional
    scaling \cite{mds}, this method addresses the problem by trying to find a low
    dimensional distance matrix that is as close as possible to a multidimensional
    version that encodes the relationship between the points in the original space.
    Then the algorithm uses the founded matrix to build a correspondent configuration
    of points in a low dimensional space that will act as a less complicated
    representation of the original dataset.\\

    This work aims to develop the theorical and practical aspects of the methods
    described above, with the objective of build methods that can reduce a
    multidimensional manifold to a low dimensional representation. The first part
    of the work a shows a theorical framework. Then an implementation section will
    take place, followed by a result section. Finally, some conclusions are made.

    \section*{Theorical framework}

    In this section the theorical part of the methods is described.

    \subsection*{Linear local embedding}

    This method is based on two optimizations problems. The first one aims to
    find weights in order to build representation of each point. This
    representation is constructed as a linear combinations of its neighbors.
    The mathematical formulation of this problem is the following

        \begin{equation*}
            \begin{aligned}
               \underset{w_{ij}}{min} \quad \sum_{i=1}^n \lVert x_i - \sum_{j=1}^k w_{ij} x_j \rVert^2
            \end{aligned}
        \end{equation*}

    The solution to the problem can be found solving a the folowing system of
    linear equations.

        \begin{equation*}
            \begin{aligned}
               Gw_i &= e\\
               G = (x_i e^t -& V_i)^t (x_i e^t - V_i)\\
               V_i = [x_{v(1)}, &x_{v(2)} ... x_{v(k)}]\\
               e^t =& [1 ... 1]\\
            \end{aligned}
        \end{equation*}

    The second problem aims to find a configuration of points in a low
    dimensional space that is as similar as posible to the original dataset by
    using the weights founded in the previous step. The mathematical formulaiton
    is as follows.

        \begin{equation*}
            \begin{aligned}
               \underset{y}{min} \quad \sum_{i=1}^n \lVert y_i &- \sum_{j=1}^k w_{ij} y_j \rVert^2
            \end{aligned}
        \end{equation*}

    Finally the solution to the entire problem can be found by solving the
    folowing eigenvector problem

        \begin{equation*}
            \begin{aligned}
               M y^t &= \lambda y^t\\
               M =& (I-w)(I-w)^t
            \end{aligned}
        \end{equation*}

    \subsection*{Multidimensional scaling}

    This method is based on an optimization problem. The goal is to find a
    configuration of points in a low dimensional space with a distance matrix
    that is as similar as posible to the one that is generated from the original
    dataset. The mathematical formulation is as follows.

        \begin{equation*}
            \begin{aligned}
                \underset{y}{min} \quad &\lVert D^{(x)} - D^{(y)} \rVert ^2\\
                d_{ij}^{(x)} =& \lVert x_i - x_j \rVert\\
                d_{ij}^{(y)} =& \lVert y_i - y_j \rVert
            \end{aligned}
        \end{equation*}

    Finally the solution to the problem can be founded by using the spectral
    decomposition and defining $\hat{\Lambda}$ as the top values of $\Lambda$.

        \begin{equation*}
            \begin{aligned}
                K = -\frac{1}{2} H D^{(x)} H &= X^t X = V \Lambda V^t\\
                H = I - \frac{1}{n} e e^t &\quad e^t = [1 ... 1]\\
                Y =& \hat{\Lambda}^{\frac{1}{2}} V^t\\
            \end{aligned}
        \end{equation*}

    \section*{Implementation}

    In this section a seudo-code for the two methods described previously is
    shown. These algorithms were implemented using MATLAB.

    \subsection*{Local linear embedding}

    $n$ = number of points\\
    $p$ = output dimensions\\
    $k$ = number of neighbors\\

    $W$ = zeros($n$)\\
    \For{i in 1:n}{
        $x_i$ = data($i$)\\
        $k_i$ = get\_nearest\_neighbors($x_i$, $k$)\\
        $V_i$ = data($k_i$)\\
        $G$ = $(x_i e^t - V_i)^t (x_i e^t - V_i)$\\
        solve($G w_i = e$)\\
        $W(i)$ = $w_i$ / sum($w_i$)
    }
    \vspace{0.5cm}

    $M$ = $(I-W) * (I-W)^t$\\
    $V$ = get\_eigenvectors($M$)\\
    $Y$ = select\_smaller($p$, $V$)

    \subsection*{Multidimensional scaling}

    $n$ = number of points\\
    $p$ = output dimensions\\
    $k$ = number of neighbors\\

    $D_e$ = compute\_euclidean\_distance(data)\\
    $G$ = zeros($n$) //graph adjacency matrix\\

    \For{i in 1:n}{
        $x_i$ = data($i$)\\
        $k_i$ = get\_nearest\_neighbors($x_i$, $k$)\\
        $G(i,k_i)$ = $D_e$($i$, $k_i$)\\
        $G(k_i,i)$ = $D_e$($i$, $k_i$)\\
    }
    $D^x$ = compute\_distances($G$)\\

    $e$ = vector\_ones($n$)\\
    $H$ = $I - \frac{1}{n} e e^t$\\
    $A$ = -$\frac{1}{2}$ pow\_elementwise($D^x$,2)\\
    $B$ = $HAH$\\
    $V$ = get\_eigenvectors($B$)\\
    $\Lambda$ = get\_eigenvalues($B$)\\
    $V$, $\Lambda$ = select\_smallers($p$, $V$, $\Lambda$)\\
    $Y$ = $\Lambda^{1/2} V^t$

    \section*{Results}

    \subsection*{Result of local linear embedding}

    \begin{minipage}[b]{0.5\linewidth}
        \begin{center}
            \includegraphics[width=\linewidth]{figures/sphere.png}
            \captionof{figure}{Sphere manifold}
        \end{center}
    \end{minipage}
    %
    \begin{minipage}[b]{0.5\linewidth}
        \begin{center}
            \includegraphics[width=\linewidth]{figures/sphere_result.png}
            \captionof{figure}{Result of Lle algorithm}
        \end{center}
    \end{minipage}

    \subsection*{Result of multidimensional scaling}

    \begin{minipage}[b]{0.5\linewidth}
        \begin{center}
            \includegraphics[width=\linewidth]{figures/s_curve.png}
            \captionof{figure}{S curve manifold}
        \end{center}
    \end{minipage}
    %
    \begin{minipage}[b]{0.5\linewidth}
        \begin{center}
            \includegraphics[width=\linewidth]{figures/s_curve_result.png}
            \captionof{figure}{Result of Mds method}
        \end{center}
    \end{minipage}

    \subsection*{Results of both methods with the same manifold}

    \begin{center}
        \includegraphics[width=0.5\linewidth]{figures/swiss_roll.png}
        \captionof{figure}{Swiss roll manifold}
    \end{center}

    \begin{minipage}[b]{0.5\linewidth}
        \begin{center}
            \includegraphics[width=1.0\linewidth]{figures/swiss_roll_result_lle.png}
            \captionof{figure}{Result of Lle algorithm}
        \end{center}
    \end{minipage}
    %
    \begin{minipage}[b]{0.5\linewidth}
        \begin{center}
            \includegraphics[width=1.0\linewidth]{figures/swiss_roll_result_mds.png}
            \captionof{figure}{Result of Mds method}
        \end{center}
    \end{minipage}

    \section*{Conclusions}

    In this work a theorical and practical develop of the local linear embedding
    algorithm multidimensional scaling method was described. During the process
    it was found that althought the problems started with a quite complicated
    setup, the development of a theorical frameworks  allowed to generate great
    simplifications that reduced the original problems to just solve linear
    systems of equations and eigenvector problems. These simplifications made
    possible the creation of a relatively simple algorithm that can detect low
    dimensional manifolds in multidimensional spaces. This statement is
    supported in the results section where 3 dimensional manifolds were
    simplified to dimension 2 but preserving the intrinsic structure of the
    original dataset.

    \bibliographystyle{unsrt}
    \bibliography{poster}

\end{multicols}
\end{document}

### 2018-11-19 DimReduction/Mds/Documentation/Article_mds.tex
Adding s curve dataset to mds
Finishing Mds article

@@
     of non-linear dimensionality reduction using the multidimensional scaling
-    method and provide an algorithmic implementation easy and understandable. In
+    method and provides an algorithmic implementation easy and understandable. In
     this approach a distance matrix $D^x$ is constructed to approximate the
@@
 is being analyzed does not exploit all the dimensionality of the space where
-is defined. Due to this fact, another description of the goal that the
+it is defined. Due to this fact, another description of the goal that the
 dimensionality reduction techniques have is related to find a meaningful low
@@
 which means that they do not assume that the structure of the original dataset
-is encoded based on euclidean distance meassuares. Instead these algorithms try
+is encoded based on euclidean distance meassures. Instead these algorithms try
 to express their relationships in terms of geodesic distances, which are
@@
 multidimensional manifold to a low dimensional representation. To achieve this
-in the first part of the work a theorical framework is developed, in order to
+in the first part of the work a theorical framework is developed in order to
 understand all the aspects of the method. Then an implementation section will
@@
     development of the multidimensional scaling algorithm. This section shows a
-    detail explanation of the therical aspects of the method taking \cite{proof}
+    detail explanation of the theorical aspects of the method taking \cite{proof}
     as base point.
@@
 
-    Now, using that knowledge, the main expression can be expanded
+    Now, since this is a quadratic expression, a minimum can be found by making
+    $\Lambda$ as close as possible to $G \hat{\Lambda} G^t$. Then $G$ must be
+    an identity matrix.\\
 
-    \begin{equation*}
-        \begin{aligned}
-        Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
-        &=
-        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + (G \hat{\Lambda} G^t)^2]\\
-        =
-        &Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} G^t G \hat{\Lambda} G^t]\\
-        \end{aligned}
-    \end{equation*}
-
-    But, knowing that $Q^tQ = I = QQ^t$, the following properties can be developed
-
-    \begin{equation*}
-        \begin{aligned}
-            G^tG &= (V^tQ)^t V^tQ\\
-                 &= (Q^tV) V^tQ\\
-                 &= Q^tQ\\
-                 &= I
-        \end{aligned}
-    \end{equation*}
-
-    \begin{equation*}
-        \begin{aligned}
-            GG^t &= V^tQ (V^tQ)^t\\
-                 &= V^tQ (Q^tV)\\
-                 &= V^tV\\
-                 &= I\\ \\
-        \end{aligned}
-    \end{equation*}
-
-    So, the following transformations can be made to the problem
-
-    \begin{equation*}
-        \begin{aligned}
-        Tr[(\Lambda - G \hat{\Lambda} G^t G)^2 ]
-        &=
-        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} \hat{\Lambda} G^t]\\
-        =
-        Tr[&\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda}^2 G^t]\\
-        =
-        Tr[&\Lambda^2]- 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[G \hat{\Lambda}^2 G^t]\\
-        =
-        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2 G^tG]\\
-        =
-        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2]\\
-        =
-        Tr[&\Lambda^2] + Tr[ \hat{\Lambda}^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t]\\
-        \end{aligned}
-    \end{equation*}
-
-    Which takes us to the \textbf{DARK PART!}\\
-
-    We know that $\frac{\delta Tr[\Lambda G \hat{\Lambda} G^t]}{\delta G} = 2\Lambda G \hat{\Lambda}$\\\\
-    We also know that $\frac{\delta Tr(\Lambda^2)}{\delta G} = 0 = \frac{\delta Tr(\hat{\Lambda}^2)}{\delta G}$\\\\
-    Finally we know that the answer for a fixed $\hat{\Lambda}$ is $G=I$\\
-
-    \textbf{END OF DARK PART}\\
+%    Now, using that knowledge, the main expression can be expanded
+%
+%    \begin{equation*}
+%        \begin{aligned}
+%        Tr[(\Lambda - G \hat{\Lambda} G^t)^2]
+%        &=
+%        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + (G \hat{\Lambda} G^t)^2]\\
+%        =
+%        &Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} G^t G \hat{\Lambda} G^t]\\
+%        \end{aligned}
+%    \end{equation*}
+%
+%    But, knowing that $Q^tQ = I = QQ^t$, the following properties can be developed
+%
+%    \begin{equation*}
+%        \begin{aligned}
+%            G^tG &= (V^tQ)^t V^tQ\\
+%                 &= (Q^tV) V^tQ\\
+%                 &= Q^tQ\\
+%                 &= I
+%        \end{aligned}
+%    \end{equation*}
+%
+%    \begin{equation*}
+%        \begin{aligned}
+%            GG^t &= V^tQ (V^tQ)^t\\
+%                 &= V^tQ (Q^tV)\\
+%                 &= V^tV\\
+%                 &= I\\ \\
+%        \end{aligned}
+%    \end{equation*}
+%
+%    So, the following transformations can be made to the problem
+%
+%    \begin{equation*}
+%        \begin{aligned}
+%        Tr[(\Lambda - G \hat{\Lambda} G^t G)^2 ]
+%        &=
+%        Tr[\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda} \hat{\Lambda} G^t]\\
+%        =
+%        Tr[&\Lambda^2 - 2 \Lambda G \hat{\Lambda} G^t + G \hat{\Lambda}^2 G^t]\\
+%        =
+%        Tr[&\Lambda^2]- 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[G \hat{\Lambda}^2 G^t]\\
+%        =
+%        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2 G^tG]\\
+%        =
+%        Tr[&\Lambda^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t] + Tr[ \hat{\Lambda}^2]\\
+%        =
+%        Tr[&\Lambda^2] + Tr[ \hat{\Lambda}^2] - 2 Tr[\Lambda G \hat{\Lambda} G^t]\\
+%        \end{aligned}
+%    \end{equation*}
+%
+%    Which takes us to the \textbf{DARK PART!}\\
+%
+%    We know that $\frac{\delta Tr[\Lambda G \hat{\Lambda} G^t]}{\delta G} = 2\Lambda G \hat{\Lambda}$\\\\
+%    We also know that $\frac{\delta Tr(\Lambda^2)}{\delta G} = 0 = \frac{\delta Tr(\hat{\Lambda}^2)}{\delta G}$\\\\
+%    Finally we know that the answer for a fixed $\hat{\Lambda}$ is $G=I$\\
+%
+%    \textbf{END OF DARK PART}\\
 
@@
     To minimize the resulting expression, the first $p$ eigenvalues of $\Lambda$
-    can be used to define $\hat{\Lambda}$. And finally the solution to the
+    can be used to define $\hat{\Lambda}$. Finally, the solution to the
     problem can be found by the following procedure
@@
 
-    which implies that $Y = \hat{\Lambda}^{1/2} V^t$. Which solves the problem.
+    which implies that $Y = \hat{\Lambda}^{1/2} V^t$. And that solves the problem.
 
@@
 
-    To solve the problem the first step is rewrite each entry of matrix $D$ in
+    To solve this problem the first step is rewrite each entry of matrix $D$ in
     a vector form
@@
 a swiss roll dataset that was used as the multidimensional manifold. Figure
-\ref{swiss_res} shows the result of the devolped method.
+\ref{swiss_res} shows the result of the devolped method. Finally in Figure       
+\ref{s_curve} there is a s curve dataset and Figure \ref{s_curve_res} shows its  
+corresponding simplified output.
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{images/swiss_roll_result.png}
+    \includegraphics[width=0.6\linewidth]{images/swiss_roll_result.png}
     \caption{Swiss roll dataset reduced}
     \label{swiss_res}
+\end{figure}
+
+\begin{figure}[H]                                                                
+    \centering                                                                   
+    \includegraphics[width=0.9\linewidth]{images/s_curve.png}                    
+    \caption{S curve dataset}                                                    
+    \label{s_curve}                                                              
+\end{figure}
+
+\begin{figure}[H]                                                                
+    \centering                                                                   
+    \includegraphics[width=0.9\linewidth]{images/s_curve_result.png}             
+    \caption{S curve dataset reduced}                                            
+    \label{s_curve_res}                                                          
 \end{figure}
@@
 In this work a theorical and practical development of the multidimensional
-scaling method was permed. During the process it was found that what seemed like
+scaling method was conducted. During the process it was found that what seemed like
 a fairly complicated problem, can be simplified to simple matrix operations
@@
 framework, the construction of a relatively simple algorithm that can detect
-low dimensional manifolds embedded in multidimensional spaces was possible.
-Finally the results shown that this technique has the ability to make
+low dimensional manifolds embedded in multidimensional spaces became possible.
+The results shown that this technique has the ability to make
 contributions to the analysis proces of multidimensional datasets by
-simplifying its features but preserving the intrinsic structure.
+simplifying its features but preserving the intrinsic structure. Finally, as
+future work, this method can be improved by creating new ways to compute the
+matrix $D^x$. It would be interesting to implement this method using GPU
+computations to improve its performance and use it as a mechanism of
+data preprocessing in machine learning algorithms.
 
@@
 
-    And if $Q$ is also sqare, then
+    And if $Q$ is also square, then
 
@@
 
-    So, if a new matrix $\Lambda$ is define as a diagonal matrix that contains
+    So, if a new matrix $\Lambda$ is defined as a diagonal matrix that contains
     the eigenvalues of $A$, then the computation can be expressed as

### 2021-04-10 Dnn/Python/Execute.py
Adding neural net implemented in python

@@ -1,0 +1,24 @@
+from NNet import NNet
+from numpy import array
+
+arch = [2, 2, 2]
+input_data = [0.05, 0.1]
+labels = [0.01, 0.99]
+eta = 0.5
+
+weights = [ [], [] ]
+weights[0] = [ [0.15, 0.25],
+               [0.20, 0.30],
+               [0.35, 0.35] ] # Esta ultima linea son los pesos de los bias
+
+weights[1] = [ [0.40, 0.50],
+               [0.45, 0.55],
+               [0.60, 0.60] ] # Esta ultima linea son los pesos de los bias
+weights = array( weights )
+
+nnet = NNet(arch)
+
+for i in range(10):
+    error, weights = nnet.train(input_data, labels, weights, eta)
+    output = nnet.predict(input_data, weights)
+    print(error, output)

### 2021-04-10 Dnn/Python/NNet.py

@@ -1,0 +1,137 @@
+import numpy as np
+
+# Ejemplo de una red neuronal simple
+# https://mattmazur.com/2015/03/17/a-step-by-step-backpropagation-example/
+
+from Functions import sigmoid, d_sigmoid, mse, d_mse, add_bias # importar las funciones con sus respectivas derivadas
+                                                               # add_bias esta explicado en el archivo de funciones
+class NNet:
+    # Este es el constructor de la clase, recibe una lista que contenga el numero
+    # de neuronas que se quieren en cada capa sin tener en cuenta el bias.
+    # Es decir, para una red que tenga 3 entradas, 2 capas ocultas con 5 y 4
+    # neuronas respectivamente y 1 sola salida, arch tendria que ser [3, 5, 4, 1]
+    def __init__(self, arch):
+        self.arch = arch
+        self.n_layers = len(arch)-1 # se le tiene que restar 1 porque las listas
+                                    # empiezan en 0, entonces si la red neuronal
+                                    # tiene entrada, 2 capas y salida, la lista
+                                    # de arquitectura tendria 4 posiciones,
+                                    # de 0 a 3
+
+    # Esta funcion hace el papel de forward, recive la entrada de la red (data)
+    # y los pesos de cada capa. El arreglo de los pesos debe tener el mismo
+    # numero de elementos que la lista de arquitectura (arch) menos uno. Esto
+    # debido a que los pesos conectan las capas. Es decir, si la red es la del
+    # ejemplo, entonces existe una posicion que representa los pesos de la
+    # conexion de la entrada a la primera capa oculta, y otra que representa
+    # los pesos de la capa ocula a la salida.
+
+    # el formato de data es simplemente una lista que contanga una posicion por
+    # cada entrada a la red.
+
+    # el formato de los pesos es: en las filas van los pesos que estan relacionados
+    # con las neuronas de las que sale la informacion y en las columnas estan
+    # a las neuronas a las que llegan. Es decir, el primer indice de la lista
+    # de pesos indica con cual conjunto de pesos se va a trabajar, weights[0]
+    # es una matriz que tiene los pesos que conectan la entrada con la primera
+    # capa oculta. Luego en esa matriz la posicion 0,2 siginifica el peso que va
+    # de la entrada 0 a la neurona 2 de la capa oculta (recuerda que los indices
+    # empiezan en 0, entonces la neurona de la posicion 2 es la tercera de esa
[…]
